### 2018-04-12 docs/methods/research_highlights.tex
NM

\documentclass[11pt]{article}

\begin{document}
\begin{center}
  {\bf Research Highlights} \\
  A boundary-integral framework to simulate \\ viscous erosion of a
  porous medium \\
  Bryan D.~Quaife and M.~Nicholas J.~Moore
\end{center}

\begin{itemize}
  \item A highly accurate numerical method for the simulation of erosion
    in a two-dimensional porous medium that resolves multibody
    interactions
  \item A method to allow for erosion while artificially fixing the area
    so that numerical and analytical results can be compared
  \item A demonstration of channelization occurring in porous medium and
    the resulting drag and resistivity of the flow

\end{itemize}

\end{document}