### 2018-07-03 docs/methods/revisions.tex
Addressed Reviewer 1

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{amsmath,amsfonts,amssymb,stmaryrd}
\usepackage{color}
\newcommand{\comment}[1]{{\color{blue} #1}}

\begin{document}

\noindent
We thank the reviewers for their reviews.  Below is a summary of the
changes.  The revised manuscript also addresses a typo that was in
equation (10).

\section*{Reviewer 1}

\noindent
\comment{In section 3.1, formula 29 is missing a $\Delta s_{j}$}
\begin{itemize}
  \item Fixed
\end{itemize}

\noindent
\comment{In section 4.1, the symbols $v$ and $u$ haven't been defined
previously}
\begin{itemize}
  \item $\mathbf{u} = (u,v)$ is now defined leading up to Section 4.1.
\end{itemize}

\section*{Reviewer 2}
\noindent
\comment{The time scale of erosion totally disappears from the analysis.
How would this time scale be set in practise?}
\begin{itemize}
  \item After equation (1), the erosion timescale is described.
\end{itemize}

\noindent
\comment{On page 23: the scaling of Fd with 4*pi*U is not dimensionless.
It seems to me like the drag coefficient might be a better quantity to
base the discussion on.}
\begin{itemize}
  \item 
\end{itemize}

\noindent
\comment{The aft and rear stagnation points both recede eventhough they
are subject to zero shear. Is it just a consequence of the
regularization?}
\begin{itemize}
  \item 
\end{itemize}

\end{document}