### 2019-04-22 docs/proposals/ErosionWhitePaper.tex
working on DOE white paper

\documentclass[11pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage[sort,nocompress]{cite}
\usepackage{color}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{lastpage}
\usepackage{pgfplots}
\usepackage{rotating}
\usepackage{siunitx}
\usepackage{subfigure}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usepackage{titlesec}
\usepackage{todonotes}
\usepackage{wrapfig}
\usepackage{verbatim}
%\usepackage{showkeys}

\usetikzlibrary{shapes,arrows}

\addtolength{\oddsidemargin}{-0.75in}
\addtolength{\evensidemargin}{-0.75in}
\addtolength{\textwidth}{1.5in}
\addtolength{\topmargin}{-0.75in}
\addtolength{\textheight}{1.5in}
% For 11pt size

\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}

\newcommand{\bd}{\partial}
\newcommand{\BL}{\mathrm{BL}}
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\cc}{\mathbf{c}}
\newcommand{\dlp}{{\mathrm{dlp}}}
\newcommand{\ee}{\mathbf{e}}
\newcommand{\ff}{\mathbf{f}}
\newcommand{\FF}{\mathbf{F}}
\renewcommand{\gg}{\mathbf{g}}
\newcommand{\II}{\mathcal{I}}
\newcommand{\mcaption}[2]{\caption{\small \em #1}\label{#2}}
\newcommand{\nn}{\mathbf{n}}
\newcommand{\NN}{{\mathbf{N}}}
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\pderivtwo}[2]{\frac{\partial^{2} #1}{\partial #2^{2}}}
\newcommand{\pphi}{{\boldsymbol{\phi}}}
\renewcommand{\Re}{{\textrm{Re}}}
\newcommand{\rr}{{\mathbf{r}}}
\newcommand{\RR}{{\mathbb{R}}}
\newcommand{\slp}{{\mathrm{slp}}}
\newcommand{\ssigma}{{\boldsymbol{\sigma}}}
\newcommand{\SLP}{{\mathtt{SLP}}}
\newcommand{\DD}{{\mathcal{D}}}
\renewcommand{\SS}{{\mathcal{S}}}
\newcommand{\ttau}{{\boldsymbol{\tau}}}
\newcommand{\uu}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\VV}{{\mathcal{V}}}
\newcommand{\xx}{\mathbf{x}}
\newcommand{\yy}{\mathbf{y}}
\newcommand{\thL}{$\theta$--$L$}

\pagestyle{fancy}
\lhead{\footnotesize Bryan Quaife}
\chead{\footnotesize Viscous Erosion of a Porous Medium}
\rhead{\footnotesize \thepage}
\cfoot{}

%\usepgfplotslibrary{external}
%\tikzexternalize

\begin{document}
\begin{center}
Viscous Erosion of a Porous Medium \\
Bryan Quaife, Assistant Professor \\
Florida State University \\
512-436-1148, bquaife@fsu.edu \\
DOE/Office of Science Program Office: ASCR-Applied Mathematics \\
Funding Opportunity Announcement Number: DE-FOA-0001968
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{wrapfigure}[24]{r}{0.7\textwidth}
\centering
\includegraphics[width=0.68\textwidth]{figs/50bod.pdf}
\caption{\label{fig:50bod} \em Simulation of 50 bodies eroding in Stokes
flow under the action of fluid-mechanical erosion. The flow is
horizontal (left to right) and the 6 snapshots are evenly spaced in
time.  The local erosion rate on the bodies is proportional to the
magnitude of the vorticity which we compute in the fluid bulk (color
scale).  Erosion not only diminishes the size of the bodies but also
alters their shapes considerably. A few well-defined channels develop in
the space between bodies.}
\end{wrapfigure}

Flow-induced erosion deteriorates and reshapes solid material over a
range of scales found in nature, including massive land formations
(mountains undergoing wind erosion) and watercourses (rivers and streams
undergoing hydraulic erosion).  Though less visible, erosion is working
at the very smallest scales, slowly deteriorating the individual
constituents of porous media (e.g.~soil, sand, or clay) or biological
structures like plaque and biofilms. Motivated by such examples, we are
interested in studying fluid-mechanical erosion in Stokes flow---the
most relevant regime for groundwater and bio-fluid applications.

In our recent publication~\cite{qua-moo2018}, we describe a suite of
numerical methods to simulate erosion in a viscous fluid
(Figure~\ref{fig:50bod}).  The highlights of our numerical methods are:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item {\bf Boundary Integral Equation Fluid Solver}: The
  incompressible Stokes equations are solved with a boundary integral
  equation (BIE).  By using a BIE, the complex geometries are naturally
  handled, high-order accurate discretizations are deployed, and fast
  summation methods result in optimal complexity.

  \item {\bf Interface Tracking}: Instead of tracking the $(x,y)$
  coordinates of the of the eroding interfaces, we track the {\thL}
  coordinates, where $\theta$ is a local tangent angle, and $L$ is the
  total interface length.  The governing equations of the {\thL}
  coordinates offers several advantages in the ability to numerically
  stabilize the evolving interface.

  \item {\bf Regularizations of the Interface}: Erosion results in
  numerically challenging corners of the bodies at stagnation points of
  the flow.  We use a regularization term and a smoothing operation to
  control the smoothness of the moving interface.

  \item {\bf Analysis of Single Eroding Bodies}: We validate our method
  by comparing numerical simulations with scaling laws.
\end{enumerate}

\section{Proposed Work}
Having developed and validated numerical methods for simulating viscous
erosion, there are two main directions that our research will follow.
First, additional numerical tools are required to simulate more
realistic scenarios such as denser packings, transport, and
three-dimensional erosion.  Second, we will investigate physically
motivated questions regarding the geometry and flow inside an eroding
porous medium.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved Numerical Methods}
\begin{wrapfigure}[22]{r}{0.4\textwidth}
\centering
\input{figs/vort1.tikz}
\caption{\label{fig:vort1} \em The vorticity a vertical slice cutting
through a single body.  There are solid walls at $y = \pm 1$, and an
eroding body at $y = (-0.5,0.5)$.  The trapezoid rule gives large errors
near the solid walls, while the Barycentric quadrature formula results
in a smooth vorticity field.}
\end{wrapfigure}

In its current formulation, our bodies are fixed in space.  However, in
reality, eroding bodies are also transported by the flow.  To allow the
bodies to move, there are two additional tools that we will develop.
The PI is experienced with simulating moving interfaces in Stokes flow.
For example, he has developed high-order adaptive time stepping methods
for simulating deformable bodies (vesicles) in a viscous
fluid~\cite{qua-bir2015b, qua-bir2014, qua-bir2016}. More recently, the
PI and his former Ph.D.~student developed time stepping methods for
rigid body transport in a viscous fluid~\cite{bys-sha-qua2018}.  This
work extends a contact algorithm, originally developed by Lu et
al.~\cite{lu-rah-zor2017}, to accuratly simulate the dynamics of
nearly-touching rigid bodies.  The contact algorithm resolves errors
introduced by time stepping, but quadrature error introduced by the
integral equation formulation must also be addressed.  A postdoc in the
PI's group is extending a Barycentric near-singular integration scheme
developed by Barnett et al.~\cite{bar-wu-vee2015} to develop quadrature
formula to compute the shear stress and vorticity.  In
Figure~\ref{fig:vort1}, the vorticity is computed along a vertical slice
that cuts through a single eroding body.  The trapezoid rule is accurate
far from the solid walls ($y = \pm 1$) and the boundary of the eroding
body ($y = \pm 0.5$), but introduces large errors near the walls and
body.  However, the Barycentric quadrature rule is uniformly accurate
throughout the entire fluid domain.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Physics-Based Understanding of Viscous Erosion}
\todo[inline]{Characterizing anisotropy}
\todo[inline]{Coarse-grained transport (anomalous diffusion)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Biographical Information}
Since 2015, {\bf Bryan Quaife} has been an Assistant Professor in the
Department of Scientific Computing at Florida State University.  He is
also a Research Associate in FSU's Geophysical Fluid Dynamics Institute.
Dr.~Quaife received his Ph.D.~in 2011 in Applied and Computational
Mathematics at Simon Fraser University.  He was a postdoctoral fellow
for four years in Dr.~George Biros' group at the Institute for
Computational Science and Engineering at the University of Texas.  His
research interests include the development of  high-fidelity and
efficient numerical methods for simulating fluid dynamics in complex
geometries.

The co-PI of this project is {\bf Nick Moore}.  Since 2014, Dr.~Moore
has been an Assistant Professor in the Department of Mathematics at
Florida State University.  He is also a Research Associate in FSU's
Geophysical Fluid Dynamics Institute. \todo[inline]{A bit more}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Expected Budget}
\todo[inline]{I can have Michelle put one together.  Funds are needed
for the following}
\begin{itemize}
  \item Summer salary (1 month a year for 3 years)
  \item 2 RA (1 in Math, 1 on DSC)
  \item 1 joint postdoc
  \item Travel
  \item Computers
  \item FSU overhead
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\footnotesize{\bibliography{refs}}

\end{document}