### 2019-10-05 docs/transport/erosion_transport.tex
working on edits

\documentclass[preprint,10pt]{elsarticle}

\input{preambles.tex}
%\usepackage{showkeys}
\begin{document}

\title{Viscous Transport in Eroding Porous Media}

% OTHER TITLE POSSIBILITIES
% ???

\author[SH]{Shang-Huan Chiu}
\author[Nick]{M.~N.~J.~Moore}
\author[Bryan]{Bryan D.~Quaife}

\address[SH]{Department of Scientific Computing, Florida State
University, Tallahassee, FL, 32306.}
\address[Nick]{Department of Mathematics and Geophysical Fluid Dynamics
Institute, Florida State University, Tallahassee, FL, 32306.}
\address[Bryan]{Department of Scientific Computing and Geophysical Fluid
Dynamics Institute, Florida State University, Tallahassee, FL, 32306.}

\begin{abstract} 
  We extend our previous work, [{\em Journal of Computational Physics},
  {\bf 375}:1--21, 2018], to simulate fluid-mechanical erosion of
  two-dimensional porous media with nearly-touching grains.  The method
  continues to use a boundary integral equation (BIE) method to solve
  the fluid equations, and a stable second-order time stepping method to
  erode the grains.  In our previous work, we required eroding grains
  that were sufficiently separated to guarantee that the trapezoid rule
  provides a sufficient amount of accuracy. We now remove this
  restriction by introducing a Barycentric quadrature rule whose error
  is bounded independent of the geometry.  The Barycentric quadrature
  rule is computationally expensive, so it is only used when necessary,
  and the spectrally accurate fast-multipole-accelerated trapezoid rule
  is used otherwise. We also simulate and analyze transport through
  eroded porous media.  By using the Barycentric quadrature rule, the
  analysis examines flow both in the fast and slow regions with
  high-order accuracy. Our results reveal several interesting  effects
  of erosion on the tortuosity, dispersion, and pore size distributions.
\end{abstract}

\begin{keyword}
  Porous media \sep Stokes flow \sep Erosion \sep Quadrature \sep
  Transport \sep Tortuosity \sep Anomalous dispersion \sep Pore size
  distributions
\end{keyword}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Porous media flow plays an important role in many environmental and
industrial applications.  Depending on the application, length scales
can vary from $10^{-6}\mathrm{m}$ to
$10^{-1}\mathrm{m}$~\cite{mil-chr-imh-mcb-ped1998} and velocity scales
can be as small as $10^{-1}\mathrm{m/day}$~\cite{kut-scr-dav-ham1995}.
Moreover, for a single porous geometry, the pore sizes and velocities
can range over several orders of magnitude.  Numerical methods that
resolve this range of scales offer the ability to: (i) characterize
dispersion~\cite{saf1959}, (ii) quantify
mixing~\cite{leb-den-dav-bol-car-dec-bou2011, den-leb-eng-bij2011}, and
(iii) develop meaningful constitutive relationships that link the
microscopic and macroscopic realms~\cite{mil-chr-imh-mcb-ped1998}.
Examples of coarse-grained models for porous media flow include
pressure-saturation-permeability
relationships~\cite{mil-chr-imh-mcb-ped1998}, permeability-porosity
relationships~\cite{dar-mcc1998, car1937, won-kop-tom1984}, first
passage time distributions~\cite{ber-sch-sil2000, hym-den-hag-kan2019,
cve-che-wen1996}, tortuosities~\cite{hak-com-den2019, mat-kha-koz2008,
dud-koz-mat2011, kop-kat-tim1996}, concentration
distributions~\cite{ica-den2019, bel-sal-rin1992}, pore and flow
distributions~\cite{ali-par-wei-bre2017}, geometry
connectivity~\cite{knu-car2005, wes-blo-gra2001}, anomalous
dispersion~\cite{dea-qua-bir-jua2018, den-cor-sch-ber2004,
sie-ili-pri-riv-gua2019}, and more.

Flow in porous media is further complicated when boundaries dynamically
evolve in response to the fluid flow. This coupling between the geometry
and the flow occurs, for example, in applications involving
melting~\cite{bec-vis1998, rycroft2016asymmetric, jambon2018singular,
favier2019rayleigh, morrow2019moving},
dissolution~\cite{kan-zha-che-he2002, mac2015shape, moo2017,
wykes2018self}, deposition~\cite{joh-eli1995, hewett2018modelling},
biofilm growth~\cite{tan-val-wer2015}, and crack
formation~\cite{cho2019crack}. We focus on erosion, a fluid-mechanical
process that is prevalent in many geophysical, hydrological, and
industrial applications~\cite{berhanu2012shape, hewett2017evolution,
lachaussee2018competitive, lopez2018cfd, allen2019sde, amin2019role}.
When a porous geometry erodes, certain qualitative characteristics are
unveiled that affect transport through the geometry.  For example, an
eroded geometry may contain channels of high porosity, which, though few
in number and modest in volume fraction, transmit a large portion of the
flux \cite{qua-moo2018}. This arrangement results in velocities that
vary over several orders of magnitude~\cite{all-hea-lab-rei2002}.
Moreover, channelization creates heterogeneous and anisotropic medium
properties, which affect the transport of tracers such as
contaminants~\cite{cve-che-wen1996, dag1987, kon-bre1978} and
heat~\cite{nil-sto1990, ree-sto1995}.

This study consists of two main undertakings: first, high-fidelity
simulations of eroding porous media, and, second, characterization of
tracer transport through the resulting eroded geometries. Our modeling
efforts build on previous work~\cite{ris-moo-chi-she-zha2012,
moore2013self, moo2017}, in particular recent numerical methods
developed to simulate erosion in the Stokes-flow
regime~\cite{qua-moo2018}. We, however, make key improvements to the
numerical methods from~\cite{qua-moo2018} to enable simulations of more
realistic, dense suspensions of eroding grains
(Figures~\ref{fig:Eroding50vel} and~\ref{fig:Eroding50vort}).  Then, to
characterize transport through these configurations, we examine
coarse-grained variables through statistical analysis of tracer
trajectories.

\begin{figure}[H]
\begin{center}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b1}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b80}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b160}\\
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b240}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b320}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b400}
\caption{\label{fig:Eroding50vel} Six evenly spaced snapshots of 50
eroding bodies.  The background velocity is a Hagen-Poiseuille flow with
a varying flow strength that corresponds to a constant pressure drop
across the channel. The color is the magnitude of the velocity in a
logarithmic scale. The flow velocity varies over several orders of
magnitude, the grains form irregular shapes with large aspect ratios,
and the geometry becomes channelized and anisotropic.}
\end{center}
\end{figure}

Owing to the scales present in groundwater flow~\cite{bea1972}, we model
the hydrodynamics with the two-dimensional incompressible Stokes
equations.  Meanwhile, individual grains erode at a rate proportional to
the hydrodynamic shear stress~\cite{wan-fel2004,
ris-moo-chi-she-zha2012, moore2013self, par-izu2000}.  Since the
vorticity and the shear stress are equivalent on solid boundaries, we
also compute the fluid vorticity for visualization purposes
(Figure~\ref{fig:Eroding50vort}).  A variety of numerical methods for
the fluid solver are available including lattice-gas or lattice
Boltzmann~\cite{kop-kat-tim1996, dar-mcc1998}, finite
volumes~\cite{suo-liu-gan2019, den-ica-hid2018,
sie-ili-pri-riv-gua2019}, finite differences~\cite{leb-ded-dav-bou2007,
knu-car2005}, and cellular automata~\cite{rot1988}.  However, simulating
erosion requires a method that is high-order, can resolve
nearly-touching arbitrarily shaped grains, and can accurately compute
derivatives of the velocity field.  Since the governing equations are
linear and homogeneous, we address these challenges by converting the
fluid equations to a boundary integral equation (BIE).

To compute stable simulations of erosion, we use high-order methods in
both space and time. The time integration method is unchanged from
previous work~\cite{qua-moo2018}.  We apply a mild regularization and a
smoothing term to eliminate numerical instabilities that can be triggered by changes in
sign of the shear stress, and we use a stable second-order Runge-Kutta
method applied to the {\thL} coordinates~\cite{hou-low-she1994} of the
eroding grains.  In this work, we introduce a new quadrature method to
resolve dense suspensions.  The accuracy of the
trapezoid rule, which was used in previous work, is adequate for
bodies that are sufficiently separated~\cite{tre-wei2014}.  However, it
loses accuracy whenever grains are in close contact.  One of the
earliest works to address quadrature methods for nearly-singular
integrands was developed by Baker and Shelley~\cite{bak-she1986}, and in
recent years, many other near-singular integration schemes have
followed~\cite{oja-tor2015, kli-tor2018, hel-oja2008a, bea-yin-wil2016,
bea-lai2001, klo-bar-gre-one2013}.  We use a Barycentric quadrature
method~\cite{bar2014, bar-wu-vee2015} since it is a non-intrusive
modification of the trapezoid rule, and the error is guaranteed to be
uniformly bounded.  We extend the original quadrature method to compute
the velocity gradient, which is needed to evaluate the shear stress and
the fluid vorticity.

\begin{figure}[H]
\begin{center}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b1}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b80}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b160}\\
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b240}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b320}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b400}
\caption{\label{fig:Eroding50vort} The same six snapshots as
Figure~\ref{fig:Eroding50vel}.  The color is the vorticity of the fluid.
Since the rate of erosion is equivalent to the magnitude of the
vorticity, erosion is fastest in the yellow and blue regions and slowest
in the green regions.}
\end{center}
\end{figure}

To characterize transport through the resulting configurations, particle
trajectories must be computed. Depending on the application, microscale
transport can be modelled as pure advection~\cite{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019},
advection-diffusion~\cite{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\cite{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume trajectories to be governed by pure advection
(ie.~no diffusion), so the particle trajectories are identical to the
streamlines. We compute trajectories $\ss(t)$ that are initialized at
$\ss_0$ by solving the advection equation
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
  \label{eqn:tracers}
\end{align}
where $\uu$ is the fluid velocity.  Since there is no stiff diffusive
term in~\eqref{eqn:tracers}, we use a fourth-order explicit Runge-Kutta
time stepping method.  In addition, the Barycentric quadrature rule is
used to accurately compute trajectories that are close to an eroding
grain. We note that an alternative approach to compute trajectories is
to form a distribution of velocities $\uu$ and use a continuous time
random walk~\cite{mon-wei1965, den-cor-sch-ber2004, leb-den-car2008,
ber-cor-den-sch2006}.

Once the streamlines are computed, we characterize transport by
analyzing three different metrics: the tortuosity, the anomalous
dispersion, and the pore size distribution.  The local tortuosity of a
streamline that connects the inlet to the outlet is defined as the
streamline's length normalized by the linear inlet-to-outlet distance.
In porous media, the local tortuosity can be greater than
1.5~\cite{kop-kat-tim1996, mat-kha-koz2008} or even
2~\cite{dud-koz-mat2011}, depending on several factors such as the
porosity. The tortuosity of a geometry is defined by averaging the local
tortuosity over all streamlines initialized at the inlet, and the
geometry's tortuosity characterizes average particle
motions~\cite{hak-com-den2019}.  To characterize spreading, the fluid
dispersion is computed by calculating the variance of the streamline
lengths. In porous media, this spreading is often
super-dispersive~\cite{kan-dea-nun-bij-blu-jua2014, cus-hu-den1995,
dea-leb-den-tar-bol-dav2013}. Since anomalous dispersion results from
streamlines spending time in both the high and low velocity
regimes~\cite{ber-sch2001}, it is crucial to accurately resolve
streamlines near grain boundaries, as achieved in this work.  Finally,
we construct the pore-size distribution throughout the erosion process.
These distributions are required to quantify velocity
distributions~\cite{ali-par-wei-bre2017, dea-qua-bir-jua2018},
channelization~\cite{sie-ili-pri-riv-gua2019},
connectivity~\cite{knu-car2005, wes-blo-gra2001}, and to develop network
models~\cite{bry-kin-mel1993, bry-mel-cad1993, bij-blu2006}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
Our first contribution is extending the Barycentric quadrature rule of
Barnett et al.~\cite{bar-wu-vee2015} to accurately compute velocity
gradients associated with Stokes flow. This new quadrature rule permits
accurate and stable simulations of erosion, and, moreover, enables
visualization of the vorticity field throughout.

Our second contribution is computing and analyzing trajectories in
eroded geometries.  These trajectories are formed by combining the
Barycentric quadrature rule and a fourth-order Runge-Kutta method. Then,
they are used to compute tortuosities and dispersion rates of eroded
geometries. Finally, we examine the effect of erosion on the pore size
distribution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Limitations}
As in our previous work~\cite{qua-moo2018}, we consider two-dimensional
immobile eroding grains. However, this work does not limit the proximity
of  grains. Compared to our previous work, we are able to simulate much
denser suspensions with a comparable number of discretization points and
without upsampling.  While we do investigate transport in this work, we
only consider passive particles that are advected by the flow.  To model
other physical processes, such as contaminant transport, we plan to
simulate a finite Peclet number advection-diffusion equation in eroded
geometries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Outline of the Paper}
In Section~\ref{sec:formulation}, we summarize the erosion model that is
described in more detail in previous work~\cite{qua-moo2018}.  In
Section~\ref{sec:DLP}, we recast all the governing equations as layer
potentials defined in both $\RR^2$ and in $\CC$.
Section~\ref{sec:transport} describes measures for characterizing the
geometry and transport. Section~\ref{sec:method} describes the numerical
methods, with special attention paid to the new quadrature method for
computing the shear stress and the vorticity.  Section~\ref{sec:results}
presents numerical examples for a variety of dense packings of bodies.
Finally, concluding remarks are made in Section~\ref{sec:conclusions}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  We only
briefly summarize the model, and a more detailed description can be
found in previous work~\cite{qua-moo2018}.  We consider flows inside a
confined geometry $\Omega$ that contains $M$ eroding bodies with
boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$.  The boundary of the
fluid domain is $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup
\gamma_M$, where $\Gamma$ is the outer boundary, taken to be a slightly
smoothed version of the boundary of $[-3,3] \times [-1,1]$.  We only
place eroding bodies in $[-1,1] \times [-1,1]$ to create a buffer region
for the flow profile imposed at the inlet to transition to the more
complex flow intervening between the bodies. On each eroding body, $\nn$
is the normal vector pointing into the body, and $\ss$ is the unit
tangent vector pointing in the counterclockwise direction.  The normal
vector on $\Gamma$ is the outward unit normal.  Neglecting inertial
forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the eroding bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity}, \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$. We simulate erosion by alternating
between solving the fluid equations and advancing the erodible
boundaries. The strength of $\UU$ is adjusted at each time step to
achieve a constant pressure drop across the channel.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Integral Equation Formulation}
\label{sec:DLP}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Since we require an accurate solution of the incompressible Stokes
equations in complex two-dimensional geometries, we reformulate the
equations as a BIE.  This has the advantage that only the
one-dimensional boundary of the domain must be discretized, and, with
appropriate quadrature formulas and fast summation methods, the result
is a high-fidelity numerical simulation with near-optimal computational
complexity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Double-Layer Potential Formulation in $\RR^2$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Applying the same approach as our previous work~\cite{qua-moo2018}, we
start with the double-layer potential 
\begin{align}
  \DDD[\eeta](\xx) = \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:velocityDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  The double-layer potential is a
solution of the incompressible Stokes equations, but it cannot represent
all the solutions. Therefore, we complete the integral equation
formulation by adding the $M$ Stokeslets and rotlets~\cite{pow-mir1987}
\begin{align}
  S[\llambda_\ell,\cc_\ell](\xx) = \frac{1}{4\pi} \left( 
    -\log \rho \II + \frac{\rr \otimes \rr}{\rho^2}
    \right)\llambda_\ell, \quad
  R[\xi_\ell,\cc_\ell](\xx) = \frac{\rr^\perp}{\rho^2} \xi_\ell,
  \quad \ell=1,\ldots,M,
\end{align}
where $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ eroding body, $\rr^\perp = (r_2,-r_1)$, and $\rho =
\|\rr\|$. Here, $\llambda_\ell$ and $\xi_\ell$ are the Stokeslet and
rotlet strengths, respectively, corresponding to the $\ell^{th}$ body.
Then, for any sufficiently smooth geometry $\Omega$, the solution of the
incompressible Stokes equation with a Dirichlet boundary condition $\ff$
is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:BIE}
\begin{alignat}{3}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx) +
    \NN_0[\eeta](\xx), \quad &&\xx \in \bd\Omega, \\
  \llambda_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell} 
    \eeta(\yy)\, ds_\yy, &&\ell = 1,\ldots,M, \\
  \xi_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell}
    (\yy - \cc_\ell)^\perp \cdot \eeta(\yy)\, ds_\yy, 
    &&\ell = 1,\ldots,M.
\end{alignat}
\end{subequations}
Here, the null space associated with the flux-free condition of $\ff$ is
addressed with  $\NN_0$ which is the integral operator with kernel
$N_0(\xx,\yy) = \nn(\xx) \otimes \nn(\yy)$, $\xx,\yy \in \Gamma$.  In
this work, $\ff$ is the prescribed velocity, which is equal to $\UU$ on
the outer wall, $\Gamma$, and equal to zero on the eroding bodies,
$\gamma_\ell$, $\ell=1,\ldots,M$.

Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, the
Stokeslets $\llambda_\ell$, and the rotlets $\xi_\ell$, we are able to
compute, for $\xx \in \Omega$, the deformation tensor, pressure, and
vorticity, defined respectively as
\begin{align}
  \label{eqn:deformationDLP} 
  \ssigma(\xx) &= \int_{\bd\Omega} \mathbf{D}_{\ssigma}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = \frac{1}{2\pi}\int_{\bd\Omega} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \nonumber \\
    & \qquad\qquad\qquad\qquad\qquad\qquad \left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
    \right) \, ds_\yy, \\
%
  \label{eqn:pressureDLP}
  p(\xx) &= \int_{\bd\Omega} D_{p}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) \nn \cdot
      \eeta(\yy)\,ds_\yy, \\
%
  \label{eqn:vorticityDLP} 
  \omega(\xx) &= \int_{\bd\Omega} D_{\omega}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi} \int_{\bd\Omega}
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4} 
      (\rr \cdot \eeta) \, ds_\yy.
\end{align}
Since the deformation tensor of the double-layer potential is
discontinuous across $\bd\Omega$, the jump term 
\begin{align}
  \frac{1}{2} \left(\pderiv{\eeta}{\ss} \cdot \ss \right) \left[
    \begin{array}{cc}
      s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
    \end{array}
  \right],
  \label{eqn:deformationJump}
\end{align}
must be added to $\ssigma(\xx)$ for $\xx \in \bd\Omega$.  Finally, the
Stokeslets and rotlets do not exhibit any jumps across $\gamma$, so the
associated components of the deformation tensor, pressure, and vorticity
are straightforward to compute~\cite{poz1992}.  Having computed the
deformation tensor on $\gamma$, the shear stress is computed using
equation~\eqref{eqn:shearStress}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In
equations~\eqref{eqn:velocityDLP},~\eqref{eqn:deformationDLP},~\eqref{eqn:pressureDLP},
and~\eqref{eqn:vorticityDLP}, the fluid velocity, deformation tensor,
pressure, and vorticity are all written as layer potentials in $\RR^2$.
However, the quadrature method we introduce in Section~\ref{sec:method}
requires a complex-valued representation of the layer potentials for the
velocity, deformation tensor, and vorticity. The first step to form a
complex representation is to write the Laplace double-layer potential as
the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \Omega,
  \label{eqn:laplaceComplex}
\end{align}
where $x = x_1 + i x_2,y = y_1 + i y_2 \in \CC$ are the complex
counterparts of $\xx = (x_1,x_2),\yy = (y_1,y_2) \in \RR^2$, and $\eta =
\eta_1 + i \eta_2$ is the complex counterpart of $\eeta =
(\eta_1,\eta_2)$. Therefore, depending on the formulation of the layer
potential, $\Omega$ is being interpreted as a subset of $\RR^2$ or
$\CC$. Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy
integral by first finding the boundary data of $v$.  Assuming that
$\Omega$ is a simply-connected interior domain, the boundary data of $v$
satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Having computed the boundary data of the
holomorphic function $v$, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

Finally, the Stokes double-layer potential~\eqref{eqn:velocityDLP} can
be written using a Laplace double-layer
potential~\eqref{eqn:laplaceComplex} and its gradients 
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\pi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\pi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
Therefore, the Stokes double-layer potential can be written as a sum of
Cauchy integrals and its first derivative~\cite{bar-wu-vee2015}
\begin{equation}
  \begin{aligned}
    u_1(x) &= \Re (v[\psi_1](x)) + \Re (v'[y\cdot\eta](x)) 
             -x_1\Re (v'[\eta_1](x)) - x_2\Re (v'[\eta_2](x)), \\
    u_2(x) &= \Re (v[\psi_2](x)) - \Im (v'[y\cdot\eta](x)) 
         +x_1\Im (v'[\eta_1](x)) + x_2\Im (v'[\eta_2](x)),
  \end{aligned}
  \label{eqn:cauchyVelocity}
\end{equation}
where $y \cdot \eta = y_1 \eta_1 + i y_2 \eta_2 \in \CC$, 
\begin{align} 
  \psi_1=(\eta_1+i\eta_2)\frac{\Re(n)}{n}, \quad
  \psi_2=(\eta_1+i\eta_2)\frac{\Im(n)}{n},
\end{align}
and $n \in \CC$ is the complex counterpart of the outward unit normal
$\nn \in \RR^2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the velocity gradient.  The deformation
tensor at $x \in \Omega$ is found by computing the derivatives of the
expressions for $u_1$ and $u_2$ in equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Re (v'[\psi_1](x)) + 
      \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
      x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Im (v'[\psi_1](x)) - 
      \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
      \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Re (v'[\psi_2](x)) - 
      \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
      x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Im (v'[\psi_2](x)) - 
      \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
      \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as in equation~\eqref{eqn:shearStress}.
The velocity gradient is also used to compute the vorticity in the fluid
bulk.  For $x \in \Omega$, the Cauchy integral representation of the
vorticity is
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Re (v'[\psi_2](x)) + \Im (v'[\psi_1](x)) 
 + \Re (v'[\eta_2](x))+ \Im (v'[\eta_1](x)), \quad x \in \Omega.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{sec:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion in porous media leads to particular phenomena such as
channelization~\cite{berhanu2012shape}, and we are interested in
characterizing transport in such geometries.  In our previous
work~\cite{qua-moo2018}, we examined the effect of erosion on the area
fraction, flow rate, and the total drag.  However, to characterize
macroscopic signatures of the transport, other quantities must be
examined.  Here, we compute the anomalous dispersion rate, the
tortuosity, and the distribution of the pore sizes.  The first two
metrics can be defined in terms of streamlines governed by the
autonomous advection equation~\eqref{eqn:tracers}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Anomalous Dispersion}
\label{sec:dispersion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The spreading of fluid in a porous media is often characterized in terms
of anomalous dispersion~\cite{kla-rad-sok2008, den-cor-sch-ber2004}.
The anomalous dispersion rate depends on the porosity and
permeability~\cite{koc-bra1988}, but is also affected by the
distribution and shape of the grains. We calculate the anomalous
dispersion rates by analyzing the streamlines governed by
equation~\eqref{eqn:tracers} in eroded geometries. Given a set of $N_p$
trajectories, we define $\lambda_j(t)$ to be the arclength of the
trajectory
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tilde{t})\|\, d\tilde{t}, 
    \quad j=1,\ldots,N_p.
\end{align}
Then, the first and second ensemble moments are
\begin{align}
  \label{eqn:moments}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2,
\end{align}
and $\sigma_\lambda$ characterizes particle dispersion.  At early times,
the particles have not explored much of the geometry, and we expect a
ballistic motion $\sigma_\lambda \sim t$.  However, as the particles
pass the grains, their trajectories are altered, and we expect that
$\sigma_\lambda \sim t^\alpha$, with $\alpha \in (0.5,1)$, indicating
that the flow is super-dispersive.

To establish an anomalous dispersion rate, the trajectories must pass
several grains so that an asymptotic regime is reached. The geometries
that we consider are too short to observe asymptotic dispersion, so we
use a reinsertion method to form longer trajectories. Similar to the
work of others~\cite{dea-qua-bir-jua2018, puy-gou-den2019}, once a
particle reaches the outlet of the porous region, it is reinserted at
the inlet.  To minimize errors caused by reinsertion, the particle is
initialized at the discretization point that has the closest velocity to
the particle's velocity at the outlet.  After a single trajectory is
formed, it has undergone a collection of reinsertions.  Then, as a
post-processing step, the trajectory is made continuous by attaching the
tail of each segment to the origin of the next segment.

We consider trajectories in an open channel so that they can be compared
to trajectories in eroded geometries with large permeability.  The
trajectory originating at $(s_1(0),s_2(0))$ is
\begin{align}
  s_1(t) = (1-s_2(0)^2)t + s_1(0), \quad
  s_2(t) = s_2(0).
\end{align}
Averaging over $s_2(0) \in [-1,1]$, the first and second ensemble
moments are
\begin{align}
  \langle \lambda \rangle (t) = \frac{2}{3}t, \quad 
    \sigma_\lambda^2(t) = \frac{4}{45}t^2.
\end{align}
In Section~\ref{sec:results}, we compare the second moment in an open
channel with the trajectories in an eroded geometry.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tortuosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The tortuosity is a dimensionless number that quantifies the amount of
twisting of streamlines. Unlike the dispersion calculations, we do not
use reinsertion to form long trajectories.  The local tortuosity is
\begin{align}
  \tau(y_0) = \frac{\lambda(y_0)}{d}.
  \label{eqn:localTort}
\end{align}
Here, the streamline originates on the inlet cross-section $x=x_0$ at
$(x_0,y_0)$, and its arclength, $\lambda(y_0)$, is calculated until the
streamline passes the parallel outlet cross-section $x = x_0 + d$.  In
this work, we consider streamlines originating at $x=-1$ and terminating
at $x=1$, so $d=2$.  However, other choices for the terminal point when
computing the tortuosity are sometime used~\cite{dud-koz-mat2011}.  The
hydraulic tortuosity is defined by taking the average over all points on
the inlet cross-section
\begin{align}
  T = \frac{1}{d}\frac{\displaystyle\int_{S}u_1(x_0,y_0)\lambda(y_0)\,dy_0}
  {\displaystyle\int_{S}u_1(x_0,y_0)\,dy_0},
  \label{eqn:tortuosity1}
\end{align} 
where $S$ is the inlet cross-section $x = -1$, $u_1(y_0)$ is the
$x$-component of the velocity at the initial point of the streamline,
and $d=2$ is the distance between the inlet and outlet.  Note that $T
\geq 1$, and $T=1$ only if no grains are present.

The definition we made for the tortuosity allows us to compute the
tortuosity in terms of an area integral.  Assuming that the flow is
incompressible and not re-entrant, meaning that all streamlines connect
the two cross-sections, the tortuosity in
equation~\eqref{eqn:tortuosity1} is equivalent to~\cite{dud-koz-mat2011}
\begin{align}
  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
           {\displaystyle\int_\Omega u_1(\yy)\, d\yy},
  \label{eqn:tortuosity2}
\end{align}
where $\Omega$ is the fluid region between the inlet and outlet
cross-sections.  Recirculation zones are possible in viscous
fluids~\cite{hig1985}, but they are very small in the examples we
consider and have a negligible effect on the tortuosity.  Since
equation~\eqref{eqn:tortuosity2} does not require the additional work of
computing particle trajectories at every time step, we use this
definition for the majority of the tortuosity calculations.  However, we
do compare the two definitions for the tortuosity at several porosities
in Section~\ref{sec:results}.

There have been efforts to relate the tortuosity, $T$, to the porosity,
$\phi$.  For example, Matyka et al.~\cite{mat-kha-koz2008} propose the
models
\begin{subequations}
  \label{eqn:tortuosityModels}
  \begin{align}
    \widehat{T}(\phi) &= \phi^{-p}, \\
    \widehat{T}(\phi) &= 1-p \log \phi, \\
    \widehat{T}(\phi) &= 1+p (1-\phi), \\
    \widehat{T}(\phi) &= (1+p (1-\phi))^2, 
  \end{align}
\end{subequations}
where $p>0$ is a fitting parameter.  In Section~\ref{sec:results}, we
compare these four models with the tortuosity of eroding porous
geometries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pore Throat Size}
\label{sec:throats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
While transport in porous media depends on the porosity, it also depends
on the placement of the grains.  In particular, grain placement affects
velocity scales~\cite{ali-par-wei-bre2017}, correlation
structures~\cite{leb-ded-dav-bou2007}, contaminant
transport~\cite{knu-car2005},
channelization~\cite{sie-ili-pri-riv-gua2019,berhanu2012shape}, and pore
network models~\cite{bry-kin-mel1993, bry-mel-cad1993, bij-blu2006}. To
characterize the grain placement, we compute distributions of pore sizes
between neighboring grains.  To define neighboring grains, we form a
Delaunay triangulation using nodes placed at the center of each eroding
grain and at a collection of points around the boundary of the porous
media.  Then, we say that two grains are neighbors if their centers
share an edge of the triangulation~\cite{dea-qua-bir-jua2018}. The pores
of an eroded geometry are illustrated in
Figure~\ref{fig:Eroding100gap}(a). We do not consider pores between
eroding bodies and the solid wall $\Gamma$, so some of the grains near
the porous region boundary only have two neighbors. Having defined the
pore sizes, we plot its distribution in
Figure~\ref{fig:Eroding100gap}(b) and compare it with the Weibull
distribution, a distribution used by others to characterize pore
sizes~\cite{ioa-cha1993}.  In Section~\ref{sec:Eroding100}, we
investigate the effect of erosion on the pore size distribution.
\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a) The pore sizes between
neighboring grains. Two grains are said to be neighbors if they share an
edge of a Delanuay triangulation with nodes at the center of each
eroding grain. (b) The distribution of the pore sizes of an eroded
geometry. The black curve is the Weibull distribution with the same
first two moments as the data.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{sec:method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo[inline]{Bryan is here}
In line with our previous work~\cite{qua-moo2018}, we use two meshes to
simulate erosion. The integral equation is solved by discretizing the
boundary of the geometry at a set of collocation points distributed
equally in arclength (Section~\ref{sec:spatialDiscretization}).
Depending on the proximity of the target point to the source points, the
quadrature rule is either the trapezoid rule or the Barycentric
quadrature rule (Section~\ref{sec:bary}).  The selection algorithm that
determines which quadrature method is applied is described in
Section~\ref{sec:fmm}.  Once the shear stress is computed, the bodies
are eroded a single time step by using a {\thL}
discretization~\cite{hou-low-she1994, moore2013self}.  The time stepping methods for
erosion and passive particles are described in Section~\ref{sec:time}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatialDiscretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Since we use a BIE formulation, we only have to discretize the
one-dimensional boundary of the domain.  Therefore, we easily resolve
complex geometries, and differentiation is performed with spectral
accuracy in Fourier space.  We discretize each eroding body $\gamma_k$
with $N_\iin$ points and discretize the outer wall $\Gamma$ with
$N_\out$ points.  The $j^{th}$ discretization point on $\Gamma$ and
$\gamma_k$ are denoted by $\yy_j^0$ and $\yy^k_j$, respectively.  The
discretization points are initially distributed evenly in arclength, and
this equispacing is maintained for the entire simulation by using the
{\thL} formulation.  In addition, we apply
regularization~\cite{qua-moo2018} to slightly smooth the corners that
inevitably develop during erosion.

Given the discretization points of $\bd\Omega$, applying the trapezoid
rule results in the collocation method for~\eqref{eqn:BIE} 
\begin{subequations}
\label{eqn:trapLinearSystem}
  \begin{alignat}{3}
  \UU_\ell &= \sum_{j=1}^{N_\out} 
    w^0_j D(\yy^0_\ell,\yy^0_j) \eeta_j^0 +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin}
    w^k_j D(\yy^0_\ell,\yy^k_j) \eeta^k_j +
  \sum_{j=1}^{N_\out} w^0_j N_0(\yy^0_\ell,\yy^0_j)\eeta_j^0 
    \nonumber \\
  &\quad+\sum_{k=1}^M S[\llambda_k,\cc_k](\yy^0_\ell) + 
  \sum_{k=1}^M R[\xi_k,\cc_k](\yy^0_\ell),
    &&\hspace{-60pt} \ell = 1,\ldots,N_\out, \\
%
  \mathbf{0} &= \sum_{j=1}^{N_\out} 
    w^0_j D(\yy^m_\ell,\yy^0_j) \eeta_j^0 +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin}
    w^k_j D(\yy^m_\ell,\yy^k_j) \eeta^k_j \nonumber \\
  &\quad+\sum_{k=1}^M S[\llambda_k,\cc_k](\yy^m_\ell) + 
  \sum_{k=1}^M R[\xi_k,\cc_k](\yy^m_\ell),
    &&\hspace{-60pt}\ell = 1,\ldots,N_\iin, \: m=1,\ldots,M, \\
%
  \llambda_m &= \frac{1}{2\pi} \sum_{j=1}^{N_\iin} w_j^m \eeta^m_j,
  \quad \xi_m = \frac{1}{2\pi} \sum_{j=1}^{N_\iin} w_j^m
    (\yy^m_j - \cc_m)^\perp \cdot \eeta^m_j,
    &&\hspace{-60pt} m=1,\ldots,M,
\end{alignat}
\end{subequations}
where $w^k_j$ are quadrature weights that depend on $N_\iin$, $N_\out$,
and the lengths of $\gamma^k$ and $\Gamma$, and $D$ is the kernel of the
Stokes double-layer potential defined in
equation~\eqref{eqn:velocityDLP}.  The linear
system~\eqref{eqn:trapLinearSystem} is a well-conditioned second-kind
integral equation that is solved iteratively with GMRES.  If the number
of discretization points is sufficiently large, then the solution
of~\eqref{eqn:trapLinearSystem} is an accurate approximation of the
density function, Stokeslets, and rotlets.   Then, if $\xx \in \Omega$
is sufficiently far from $\bd\Omega$, an approximation of the velocity,
deformation tensor, and vorticity due to the double-layer potential is
found by applying the trapezoid rule to \eqref{eqn:velocityDLP},
\eqref{eqn:deformationDLP}, and~\eqref{eqn:vorticityDLP}, respectively.
In particular,
\begin{align}
  \label{eqn:trap}
  \uu(\xx) &= \sum_{j=1}^{N_\out} w^0_j D(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j D(\xx,\yy^k_j) \eeta^k_j, \\
%
  \ssigma(\xx) &= \sum_{j=1}^{N_\out} w^0_j \mathbf{D}_\ssigma(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j \mathbf{D}_\ssigma(\xx,\yy^k_j)
  \eeta^k_j, \\
%
  w(\xx) &= \sum_{j=1}^{N_\out} w^0_j D_\omega(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j D_\omega(\xx,\yy^k_j)
  \eeta^k_j.
\end{align}
The contributions due to the Stokeslets and rotlets require no
quadrature and are easily included in the velocity, deformation tensor,
and vorticity.  Finally, to compute the shear stress, we apply the
trapezoid rule with Fourier differentiation to approximate the jump
term~\eqref{eqn:deformationJump} of the deformation tensor, and then the
tensor is applied to the normal and tangent vectors as defined in
equation~\eqref{eqn:shearStress}.

Once the velocity is computed in $\Omega$, the tortuosity can be
computed with the Eulerian velocity field~\eqref{eqn:tortuosity2}.  We
compute the velocity at $\xx_{ij} = (-1 + i\Delta x, -1 + j\Delta y)$,
$i,j=1,\ldots,N$, where $\Delta x = \Delta y = 2/N$, and the velocity at
points inside an eroding body are assigned a value of 0.  Then, the
tortuosity is approximated as
\begin{align}
  T = \frac{\sum\limits_{i,j=1}^N \|\uu(\xx_{ij})\|\Delta x \Delta y}
      {\sum\limits_{i,j=1}^N u_1(\xx_{ij}) \Delta x \Delta y}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Barycentric Quadrature Formulas}
\label{sec:bary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
While the trapezoid rule is spectrally accurate for smooth, periodic
functions~\cite{tre-wei2014}, the derivative of the integrands grows as
the target point $\xx$ approaches $\bd\Omega$ and so the error bound
depends on the location of the target point. Therefore, if the trapezoid
rule is applied when bodies are in near-contact, or when a layer
potential is evaluated at a point close to $\bd\Omega$, then the results
become unreliable and the simulation ultimately becomes unstable.  We
thus desire a quadrature method whose error bound does not depend on the
target location.

We showed in Sections~\ref{sec:DLPcomplex} and~\ref{sec:gradDLPcomplex}
that the velocity, shear stress, and vorticity of the double-layer
representation can all be written as the sum of Cauchy integrals and its
first two derivatives.  Therefore, we can accurately compute the
velocity, shear stress, and vorticity at any target location if we
develop a quadrature rules with a uniform error bounded for Cauchy
integrals and its derivatives~\eqref{eqn:cauchy}.  Ioakimidis et
al.~\cite{ioa-pap-per1991} developed quadrature rules, that we call {\em
Barycentric quadrature rules}, to compute Cauchy integrals and their
derivatives with an error bound that is independent of $x \in \CC$.
Then, Barnett et al.~\cite{bar-wu-vee2015} used these quadrature rules
to compute the Stokes double-layer potential representation of the
velocity~\eqref{eqn:velocityDLP}, and we briefly summarize their method.
Then, we build on this work to accurately compute the gradient of the
Stokes double-layer potential so that we form the shear stress and
vorticity.

We present the quadrature rules for a simply-connected interior domain
$\Omega \subset \CC$, with any point $a \in \Omega$, and we consider the
target points $x \in \Omega$ and $x \in \Omega^c$. Then, these
quadrature rules can be applied to individual components of a
multiply-connected domain to compute layer potentials of a porous media.
The method starts with an underlying quadrature rule, and we use the
spectrally accurate trapezoid rule.  Since the quadrature points are
uniformly distributed, the quadrature weights are $w_j = L/N$,
$j=1,\ldots,N$, where $L$ is the length of $\bd\Omega$.

Again, given a complex-valued density function $\eta$, the boundary data
of $v[\eta](x)$, as defined in equation~\eqref{eqn:laplaceComplex},
satisfies the Sokhotski-Plemelj jump relation~\eqref{eqn:SPrelation}.
Since the limiting boundary data of $v$ differs when considering $x \in
\Omega$ and $x \in \Omega^c$, we denote the boundary data as $v^-$ for
$x \in \Omega$, and as $v^+$ for $x \in \Omega^c$.  Rather than directly
applying the trapezoid rule to approximate $v(x)$ in
equation~\eqref{eqn:cauchy}, we start with the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^{-}(y) - v(x)}{y - x} \,dy = 0,
    \quad x \in \Omega.
\end{align}
Since the integrand is bounded for all $x \in \Omega$, we can apply the
trapezoid rule
\begin{align}
  \sum_{j=1}^{N} \frac{v^{-}(y_j) - v(x)}{y_j - x} w_j \approx 0,
\end{align}
and the error is independent of $x$.  Rearranging for $v(x)$, we have
the interior Barycentric quadrature rule
\begin{align}
  v(x) = \frac{\sum\limits_{j=1}^N \frac{v^{-}(y_j)}{y_j - x} w_j}
  {\sum\limits_{j=1}^N \frac{1}{y_j - x} w_j}, \quad x \in \Omega.
  \label{eqn:BaryvInterior}
\end{align}
Using a similar construction and letting $a$ be any point inside
$\Omega$, the exterior Barycentric quadrature rule is
\begin{align}
  v(x) = \frac{1}{x-a} 
    \frac{\sum\limits_{j=1}^N \frac{v^+(y_j)}{y_j - x}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j - a)^{-1}}{y_j - x}w_j},
    \quad x \in \Omega^c.
  \label{eqn:BaryvExterior}
\end{align}

Similar constructions can be used to derive quadrature rules
for $v'(x)$.  For $x \in \Omega$, we have the identity
\begin{align}
  \int_{\bd\Omega} \frac{v^-(y) - v(x) + (y-x)v'(x)}{(y - x)^2} = 0,
\end{align}
and the integrand is bounded for all $x \in \Omega$.  Therefore, after
applying the trapezoid rule and rearranging for $v'(x)$, we have the
interior Barycentric quadrature rule
\begin{align}
  v'(x) = \frac{\sum\limits_{j=1}^{N}
    \frac{v^{-}(y_j) - v(x)}{(y_j-x)^2} w_j}
  {\sum\limits_{j=1}^{N} \frac{1}{y_j-x} w_j}, 
  \quad x \in \Omega.
  \label{eqn:BaryvprimeInterior}
\end{align}
Using a similar construction, the exterior Barycentric quadrature rule
for the the first derivative is
\begin{align}
  v'(x) = \frac{1}{x-a} \frac{\sum\limits_{j=1}^N
    \frac{v^+_j - v(x)}{(y_j - x)^2} w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j - x} w_j},
    \quad x \in \Omega^c.
  \label{eqn:BaryvprimeExterior}
\end{align}
Note that the value $v(x)$ is required to compute $v'(x)$ for both the
interior and exterior case, and these are available using the
Barycentric quadrature rules~\eqref{eqn:BaryvInterior}
and~\eqref{eqn:BaryvExterior}.

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  The derivation is largely based on the
work of Ioakimidis et al.~\cite{ioa-pap-per1991} (see equation (2.12)).
We start with the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\pi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x}\, dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\pi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivative
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
The quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$,
and theses are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}. In Section~\ref{sec:results}, these
quadrature rules are used to form stable eroding simulations of
nearly-touching grains, and to study dynamics of the flow in regions
arbitrarily close to eroding grains.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying the Quadrature}
\label{sec:fmm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
By using the Barycentric quadrature rule, the velocity, shear stress,
and vorticity are computed with an error that is bounded independent of
the target location. However, applied directly, it requires
$\mathcal{O}(N^2)$ operations.  By using a fast summation method, such
as the fast multipole method (FMM)~\cite{gre-rok1987}, this cost can be
reduced to $\mathcal{O}(N)$ operations.  However, the Barycentric
quadrature rules require several $N$-body calculations, and they must be
done for each component of $\bd\Omega$ individually.  The resulting cost
would be prohibitive, so we introduce a hybrid method that combines the
Barycentric quadrature rule and the trapezoid rule.  Note that the
source points of the layer potential is always one of the eroding bodies
or the outer wall, but the target point can either be on another
component of $\bd\Omega$ or it can be in the fluid bulk $\Omega$.

To compute the velocity double-layer potential, we start by applying the
trapezoid rule~\eqref{eqn:trap} accelerated with the FMM.  This
calculation requires $\mathcal{O}(N)$ operations, and we call the
resulting velocity $\vv_\trap(\xx)$.  Since the trapezoid rule is
spectrally accurate, the error of $\vv_\trap(\xx)$ is small for points
sufficiently far from $\bd\Omega$, and this region depends on the number
of discretization points $N_\iin$ and $N_\out$.  However, the trapezoid
rule needs to be replaced with a more accurate quadrature rule for
points that are too close to $\bd\Omega$.  Note that since a point is
typically only close to one or two components of $\bd\Omega$, only the
contribution of these nearby bodies needs to be replaced.  Assuming that
$\xx$ is too close to $\gamma_k$, we first subtract the inaccurate
trapezoid rule approximation of the double-layer potential due to
$\gamma_k$.  Then, the Barycentric quadrature rule is used to compute
the velocity due to $\gamma_k$ with more accuracy.  That is, the final
velocity is
\begin{align}
  \vv(\xx) = \vv_\trap(\xx) - \frac{L_k}{N_k} \sum_{j=1}^{N_\iin} 
    D(\xx,\yy^k_j) \eeta(\yy_j) + \vv^k_\bary(\xx),
  \label{eqn:velocityDecomp}
\end{align}
where $\vv^k_\bary(\xx)$ is the velocity at $\xx$ resulting from
applying the Barycentric quadrature rule to the double-layer potential
due to $\gamma_k$.  This strategy naturally extends to points that are
close to $\Gamma$, and to the scenario when $\xx$ is simultaneously
close to multiple components of $\bd\Omega$.  While the term
$\vv_\trap(\xx)$ in equation~\eqref{eqn:velocityDecomp} is computed for
all target points using the FMM, the other two terms are computed with a
direct summation.  However, these terms are only required for target
points near an eroding body or the outer wall, and these points make up
only a small fraction of the total number of points.

An identical strategy is used to compute the vorticity and the
deformation tensor.  That is, the trapezoid rule is used as a first pass
to form the vorticity and deformation tensor, and then local corrections
are made to replace the inaccuracies of the trapezoid rule.  However,
since the shear stress and vorticity are only computed once per time
step, the trapezoid rule is applied with a direct summation rather than
the FMM.  Relative to the cost of computing the velocity at each
iteration of GMRES with the FMM, the additional once-per-time-step costs
to compute the vorticity and deformation tensor are minimal.

Per grid point, applying the Barycentric quadrature rules dominates the
computational cost. Therefore, it is imperative that we only apply the
Barycentric quadrature when the trapezoid rule is inaccurate.  As a rule
of thumb, the trapezoid rule due to $\gamma_k$ achieves machine epsilon
accuracy if~\cite{bar2014}
\begin{align}
  \label{eqn:TrapCutoff}
  d(\xx,\gamma_k) = \inf_{\yy \in \gamma_k} \|\xx - \yy\| > 
    5 \frac{L_k}{N_\iin}.
\end{align}
To avoid checking if all target point $\xx$
satisfy~\eqref{eqn:TrapCutoff}, we make use of the {\thL}
variables~\cite{hou-low-she1994}.  First, given two eroding bodies
$\gamma_i$ and $\gamma_j$ with lengths $L_i$ and $L_j$, and centroids
$\cc_i$ and $\cc_j$, we check if
\begin{align}
  \label{eqn:BodyCutoff}
  \|\cc_i - \cc_j\| < \frac{L_i}{2\pi} + \frac{L_j}{2\pi} + 
    \alpha_\iin \left(\frac{L_i}{N_\iin} + \frac{L_j}{N_\iin} \right),
\end{align}
where $\alpha_\iin \geq 1$ is a parameter that needs to be determined.
In this manner, rather than using an expensive all-to-all algorithm to
compute the distance between all pairs of discretization points, we
compute the distance between pairs of circle centers.  This criteria
allows us to quickly determine bodies that contain discretization points
where the Barycentric quadrature rule might need to be applied, and the
parameter $\alpha_\iin$ accounts for the fact that the bodies are being
approximated with circles.  Assuming that the two bodies $\gamma_i$ and
$\gamma_j$ satisfy condition~\eqref{eqn:BodyCutoff}, for each point $\xx
\in \gamma_j$, we check if
\begin{align}
  \label{eqn:PointsBodyCutoff}
  \|\xx - \cc_i\| < \frac{L_i}{2\pi}
+ \alpha_\iin \left(\frac{L_i}{N_\iin} \right).
\end{align}
To determine if points on $\gamma_i$ are too close to the outer wall, we
recall that the eroding bodies are all contained in $[-1,1] \times
[-1,1]$, so a target point can only be close to the lines $y = \pm 1$.
Therefore, we first check if
\begin{align}
  \left\|\cc_i - \left[
    \begin{array}{c}
      x_i \\ \pm 1
    \end{array}
    \right]
  \right\| < \frac{L_i}{2 \pi} + \alpha_\out \frac{L_\out}{N_\out},
\end{align}
where $x_i$ is the $x$-coordinate of $\cc_i$. If body $\gamma_i$
satisfies this condition, for each point $\xx=(x,y) \in \gamma_i$, we
apply the Barycentric rule to points that satisfy
\begin{align}
  \label{eqn:PointsWallCutoff}
  |y \pm 1| < \frac{L_i}{2 \pi} +\alpha_\out \frac{L_\out}{N_\out}.
\end{align}
Finally, to determine if a target point $\xx$ in the fluid bulk requires
the Barycentric quadrature rule, we only check
conditions~\eqref{eqn:PointsBodyCutoff}
and~\eqref{eqn:PointsWallCutoff}.

To determine appropriate values for $\alpha_\iin$ and $\alpha_\out$,
we fixed an eroded geometry and computed an accurate solution by using
the Barycentric quadrature rule for all discretization points.  Then,
for multiple values of $\alpha_\iin$ and $\alpha_\out$, we computed the
velocity field with the trapezoid rule for all points that do not
satisfy conditions~\eqref{eqn:PointsBodyCutoff}
and~\eqref{eqn:PointsWallCutoff}.  By comparing these two velocities, we
find that $\alpha_\iin = 4$ and $\alpha_\out = 4$ give sufficient
accuracy to maintain stability while keeping the number of points that
require the expensive Barycentric quadrature rule to a minimum.  We use
these values for all of our numerical simulations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Integration}
\label{sec:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We use the time stepping method outlined in Section 3.3 of our previous
work~\cite{qua-moo2018} which we briefly summarize here.  The erosion
rate loses differentiability if the shear stress changes sign, and this
leads to corners developing on $\gamma$ and numerical instabilities.
Therefore, we modify the erosion rate, $V_\nn$ in
equation~\eqref{eqn:erosionModel}, with
\begin{align}
  \Vn = \CE \, \abs{\tau} + \epsilon \langle\abs{\tau}\rangle \left(
    \frac{L}{2\pi} \kappa - 1 \right),
\end{align}
where $\epsilon \ll 1,$ $\langle \cdot \rangle$ is the spatial average,
$L$ is the length of $\gamma$, and $\kappa$ is the curvature of
$\gamma$.  This new term penalizations regions of high curvature, and
the normalization term guarantees that the regularization only changes
the body's shape, but not its total length.

Rather than tracking the $(x,y)$ coordinates, the {\thL} variables are
tracked. In addition, a tangential velocity field is added to each body
so that the discretization points remain equispaced in arclength for all
time.  Additionally, to increase the overall stability of the method, a
narrow Gaussian filter is applied to the erosion rate at each time step.
Time stepping is performed with a second-order Implicit-Explicit
Runge-Kutta method. In particular, the diffusive term corresponding to
the curvature penalization term is discretized implicitly, and all other
terms, which are non-stiff, are treated explicitly.  By using this time
stepping method in conjunction with the Barycentric rule, we stably
evolve the eroding bodies.

To compute the tortuosity and the anomalous dispersion rates
(Section~\ref{sec:transport}), we require the streamlines that are
governed by equation~\eqref{eqn:tracers}.  To characterize the
transport, it is crucial that the streamlines are computed accurately.
If a low-order time stepping method is used, or if $\uu(\ss(t))$ is
inaccurate, then a trajectory $\ss(t)$ can unphysically enter a grain,
rendering the trajectory meaningless to quantify transport.  However, if
we simply ignore trajectories that pass close to grains, this would
significantly skew the characterization of the transport. Therefore, we
use high-order quadrature and time stepping methods.  In particular,
depending on the proximity of $\ss(t)$ to $\bd\Omega$
(Section~\ref{sec:fmm}), we apply the trapezoid rule or the Barycentric
quadrature rule.  For time stepping, we use a fourth-order explicit
Runge-Kutta method.  Since the trajectories are computed with high-order
methods in both space and time,  we are able to simulate dynamics very
close to the eroding bodies (see Figures~\ref{fig:Eroding20tracer}
and~\ref{fig:Eroding100tort}).

Once a collection of trajectories are formed, they are used to quantify
the dispersion and the tortuosity.  We use $N_p = 1000$ streamlines so
that the statistics have converged~\cite{bel-sal-rin1992}.  However, as
described in Section~\ref{sec:dispersion}, a reinsertion method is used
to compute trajectories that are sufficiently long to observe an
asymptotic anomalous dispersion rate.  To compute the tortuosity using
the trajectories, we consider trajectories crossing between the two
cross-sections $x=-1$ and $x=1$.  Then, we discretize
equation~\eqref{eqn:tortuosity1} as
\begin{align}
  T = \frac{1}{d}\frac{\displaystyle\sum_{i=1}^{N_p} 
    u_1(y_i) \lambda(y_i) \Delta y}
  {\displaystyle\sum_{i=1}^{N_p} u_1(y_i) \Delta y}, 
\end{align}
where $\Delta y = 2/(N_p + 1)$ and $y_i = -1 + i \Delta y$,
$i=1,\ldots,N_p$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We now present numerical results on dense packings of bodies eroding in
Stokes flow. Moreover, we analyze transport through the eroded
geometries by examining coarse-grained variables.  Each eroding body is
initially a circle with center $\cc_i$ and radius $r_i$.  The center and
radius are chosen at random, and the body is accepted if it is contained
in $[-1,1] \times [-1,1]$ and is sufficiently separated from all other
bodies.  Since the error of our quadrature method is bounded independent
of the geometry, we consider bodies that are separated by less than 10\%
of an arclength spacing.  The randomized method is repeated until the
initial geometry reaches a desired initial porosity.

Once the geometry is initialized, each body has length $L_i = 2\pi r_i$,
and they are each discretized with $N_\iin$ points.  A no-slip boundary
condition is imposed on each eroding body $\gamma_i$.  For all
simulations, we discretize the outer wall $\Gamma$ with $N_\out=1024$
points.  For all but the first example, the Dirichlet boundary condition
on $\Gamma$ is a Hagen-Poiseuille flow, and the flow rate is adjusted at
each time step to maintain a constant pressure drop. Since the fluid
equations are linear, this is achieved by computing the pressure near
the inlet and outlet at each time step, and then scaling the flow rate
appropriately~\cite{qua-moo2018}.

In each of the examples, we plot the vorticity in the fluid bulk.  Since
the vorticity and the shear stress agree on surface of the eroding
bodies, the vorticity is a useful method to visualize the erosion rate.
In particular, the rate of erosion is largest in regions where the
magnitude of the vorticity is largest.

The erosion rate loses regularity at stagnation points, which inevitably
leads to corner formation on the bodies.  As described in
Section~\ref{sec:formulation} and our previous work~\cite{qua-moo2018},
we introduce a curvature penalization term and a Gaussian smoothing step
in order to ameliorate corner formation.  For all examples, we use the
smoothing parameters $\eps=15/256$ and $\sigma=10/256$, and the time
step size is $\Delta t = 10^{-4}$.

The common characteristic of each of the experiments is near-contact
between the eroding bodies, outer walls, and streamlines. We use our
numerical methods to describe and visualize the following examples:
\begin{itemize}
  \item{\bf Single Body Close to a Wall}: We consider a single eroding
  body close to the outer wall at $y=-1$.  We impose a shear flow
  centered at $y=-1$ and compare the eroding body's shape to a similar
  experiment of Spagnolie and Mitchell~\cite{mit-spa2017}.

  \item{\bf 20 Bodies at a Medium Porosity}: We consider 20 eroding
  bodies with a medium initial porosity.  After computing accurate
  streamlines, the tortuosity and anomalous dispersion rates are
  computed and compared to those of an open channel.

  \item{\bf 20 Bodies at a Low Porosity}: We consider 20 eroding bodies
  with a low initial porosity.  We examine the effect of the lower
  porosity on the tortuosity and anomalous dispersion rates.

  \item{\bf 100 Bodies at a Medium Porosity}: We consider 100 eroding
  bodies with a medium porosity.  We compute the tortuosity, anomalous
  dispersion rates, and the pore throat size distributions.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
First, consider a single eroding body close to a solid wall with the
shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
imposed on the outer wall. Mitchell and Spagnolie~\cite{mit-spa2017}
performed a similar three-dimensional experiment using a second-order
quadrature method.  They initialized the body with its center located
$1.5$ radii above the solid wall.  In two dimensions, the trapezoid rule
is spectrally accurate, and an analogous initial configuration is
possible. However, unless the number of discretization points is
increased, the trapezoid rule becomes obsolete if the eroding body is
initialized closer to the solid wall. Therefore, to simulate erosion
near a solid wall at reasonable levels of discretization, we use our
hybrid method that combines the trapezoid rule and the Barycentric
quadrature rule.

We consider an eroding body that is initially a circle of radius
$r=0.4$.  We discretize the eroding body with $N_\iin=256$
discretization points, and its center is chosen so that the distance
between the body and the solid wall is $h$, $h/2$, and $h/10$, where $h
= 2\pi r/N_\iin$.  If we used the trapezoid rule and required an error
that is comparable to the Barycentric quadrature rule, the body with an
initial distance of $h/10$ from the solid wall would require $6,400$
discretization points, and the outer wall would require $50,000$
discretization points.

In Figure~\ref{fig:NearWall}, we superimpose the eroding body's shape at
equispaced time steps.  For the three initial configurations, the shear
stress is positive for all time, but varies over several orders of
magnitude.  Therefore, we color the eroding body's boundary with the
logarithm of the shear stress.  Since the shear stress is always
positive, the erosion rate is smooth and corners do not develop, but
interesting shapes form nonetheless.  For instance, in the top half of
the body, there is a sudden increase in the shear stress, and this leads
to a region of high curvature.  This behavior is also present in three
dimensions~\cite{mit-spa2017} (Figure 7(c)).  The biggest difference
between the two- and three-dimensional results is the presence of a
recirculation zone.  In three dimensions, there is no recirculation
between the solid wall and the spherical body~\cite{cha-feu2003}, but
this is possible in two dimensions~\cite{chw-wu1975, hig1985}. To
visualize the flow, we plot the vorticity of the final time step from
Figure~\ref{fig:NearWall} in Figure~\ref{fig:NearWall_vort}.  The
recirculation is present in the region where the vorticity is smallest,
but its size is very small.

\begin{figure}[htbp]
\begin{center}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r1h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d5h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.38\textwidth}
\includegraphics[width =\textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{}
\end{subfigure}
\caption{\label{fig:NearWall} A single body eroding in a Stokes flow.
The color is the logarithm of the shear stress. Therefore, erosion is
fastest in the red regions (upper half) and slowest in the blue regions
(lower half).  The body is initialized at three different distances from
the lower wall: (a) $h$, (b) $h/2$, and (c) $h/10$.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r1h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d5h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d1h_vort}
\caption{}
\end{subfigure}
\caption{\label{fig:NearWall_vort} The vorticity of the fluid with a
single body eroding at time t=0.1. The initial distance from the body to
the solid wall are: (a) $h$, (b) $h/2$, and (c) $h/10$.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 8. Therefore, $U=1$ once all the bodies have
vanished.  Snapshots of the bodies at four equispaced times are shown in
Figure~\ref{fig:Eroding20vort}.  The color is the vorticity, whose
absolute value is proportional to the erosion rate on solid boundaries.
Initially, several of the eroding bodies are closer to the outer wall
than the $5h$ threshold required for the trapezoid rule to achieve
machine precision.  In particular, the distance between bodies 1, 6, 13,
and 15 and the outer wall is $1.3h$, $2.9h$, $2.8h$, and $1.3h$,
respectively, where $h$ is the arclength spacing of the outer wall
$\Gamma$.  In addition, the distance between several pairs of eroding
bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18, is too small to
be accurately resolved with the trapezoid rule.  By using the
Barycentric quadrature rule, the interaction between these
nearly-touching bodies can be accurately resolved to the desired
accuracy, and erosion can be simulated until all the bodies have
vanished.

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} The erosion of 20 nearly touching
bodies, each discretized with $N_\iin=256$ points, in a Hagen-Poiseuille
flow.  The four snapshots are evenly spaced in time, and the color is
the fluid vorticity. In the fourth frame, bodies 4, 12, and 17 have
vanished completely, and body 19 has almost completely eroded.}
\end{center}
\end{figure}

Erosion causes the space between nearly touching bodies to quickly
expand, and flat faces develop along the regions of near contact.  This
qualitative behavior is seen in Figure~\ref{fig:Eroding20vort} between
bodies 3 \& 4, 15 \& 16, and was also observed in previous work.
However, by resolving the interaction between bodies that are much
closer together than in previous work, we observe that very little
erosion occurs between certain pairs of bodies, at least initially.  For
instance the opening between bodies 1 \& 6, 3 \& 9, and 5 \& 13 grow
much slower than the opening between bodies 15 \& 16.  A common feature
of the pores that grow slowly is that they are nearly perpendicular to
the main flow direction, resulting in a small erosion rate.

Following our previous work~\cite{qua-moo2018}, we begin by analyzing
the effect of erosion on the area fraction and the flow rate.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  The general trend of the area fraction
resembles our previous work~\cite{qua-moo2018} (Figure 10(a)), but with
a larger initial area fraction.  In
Figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\cite{qua-moo2018} (Figure 10(b)), except that the initial flow
rate is an order of magnitude smaller because of the larger initial area
fraction.  Starting around normalized time $0.2$,
Figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as a exponential law.  The line of
best fit is
\begin{align} 
  \log_{10} U \approx 4.32 \frac{t}{t_f} + c, 
\end{align}
$t_f$ is the time required for all the bodies to erode, and $c$ is a
fitting parameter. By construction $U = 1$ when $t=t_f$, so $c=-4,32$,
and the flow rate is approximated as
\begin{align}
  U \approx \exp\left(9.94 \left(\frac{t - t_f}{t_f} \right)\right),
\end{align}
which is the dashed line in Figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/flow_rate20densen}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The area fraction of 20
nearly touching eroding bodies versus normalized time. (b) The flow
rate, $U$, for a fixed pressure drop across the channel versus
normalized time.  The flow rate is initially very small, but it
eventually increases as a power law (dashed line) towards the flow rate
$U=1$ that occurs once all the bodies have eroded.}
\end{figure}

In Figure~\ref{fig:Eroding20vort}, we observe that erosion creates a
network of channels from the inlet to outlet where the velocity and
vorticity, and therefore erosion rate, are much larger than other
regions.  These channels can be further visualized by considering the
streamlines.  In Figure~\ref{fig:Eroding20tracer}, we freeze the
geometry at the second snapshot from Figure~\ref{fig:Eroding20vort} and plot 200 streamlines that are initialized at $x=-1$ and equally
spaced in $y \in (-1,1)$.  The streamlines are shown at five different
times, and the final plot is a zoom in of the lower right quadrant of
the fifth time step, but with additional streamlines.  Since we use a
high-order quadrature rule and time stepping method, we resolve
streamlines that come very close to the eroded bodies.  There are three
clear regions where the streamlines are most concentrated, corresponding
to the regions of highest velocity.  Two of these regions are located
between the bodies and the solid walls at $y=\pm 1$, and the third cuts
through the porous region with the upper part of the channel formed by
bodies 1, 4, 6, and 8.  Since the flow is fastest in these regions, the
shearing is largest, and this causes the channels to continue to open
fastest as observed in Figure~\ref{fig:Eroding20vort}.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150n}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from Figure~\ref{fig:Eroding20vort}. The streamlines are
initialized at $x=-1$ and are equally spaced in $y \in (-1,1)$. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot, but with additional streamlines.
Since we use the Barycentric quadrature rule and the fourth-order
Runge-Kutta method, some streamlines are very close to the grains
without penetrating them.}
\end{center}
\end{figure}

Next, we use the streamlines to compute the tortuosity and the anomalous
dispersion rates of eroding geometries.  To achieve statistical
convergence, we use $N_p = 1000$ streamlines~\cite{bel-sal-rin1992}. To
compute the tortuosity, we require the velocity at the inlet $x=-1$.
These normalized velocities are plotted in
Figure~\ref{fig:Eroding20tort}(a) for the eroded geometry at porosity
$\phi = 62.9\%$ (Figure~\ref{fig:Eroding20tort}(c)).  The velocities are
similar to the work of Matyka et al.~\cite{mat-kha-koz2008} (see Figure
4(a)), except that our cross-section, by construction, does not cut
through any of the grains.  Next, in Figure~\ref{fig:Eroding20tort}(b),
we plot the local tortuosity~\eqref{eqn:localTort} by calculating the
relative length of each streamline as it traverses the channel from
$x=-1$ to $x=1$.  The local tortuosity ranges from 1 to 1.27, meaning
that one of the streamlines is 27\% longer than it would have been if
the grains were absent.  The average streamline is 9.79\% longer or
equivalently the tortuosity of the geometry is $1.098$.  Again,
comparing the local tortuosity to Matyka~\cite{mat-kha-koz2008} (Figure
4(b)), the results are qualitatively similar. However, since our initial
cross-section does not cut through the grains, the local tortuosity does
not have any gaps.  Discontinuities in local tortuosity occur when
nearby streamlines diverge to circumvent a grain.  In
Figure~\ref{fig:Eroding20tort}(c), we plot pairs of streamlines
associated with the ten largest jumps in the local tortuosity, with each
pair of corresponding streamlines plotted in the same color.

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =0.97\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity of a porous geometry
that initially had 20 grains and has eroded to a porosity of 62.9\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1,y)$,
normalized by its maximum velocity of $2.98 \times 10^{-3}$. (b) The
local tortuosity $\tau(y)$ on the cross section $x = -1$.  The spatially
averaged tortuosity through the channel is $1.098$.  (c) The streamlines
resulting in the ten largest differences between neighboring
streamlines.  Neighboring streamlines have the same color.}
\end{figure}

In Figure~\ref{fig:Eroding20Transport}(a), we plot the tortuosity as a
function of the porosity. The initial porosity is $\phi = 37.68\%$, and
the initial tortuosity is $T = 1.16$.  The tortuosity is computed with
both the length of the streamlines~\eqref{eqn:tortuosity1} (red stars)
and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in Figure~\ref{fig:Eroding20tort}(c).
The two tortuosity formulas give similar results, and any discrepancy
can be accounted for by slow regions of recirculation and from applying
quadrature to compute the tortuosity.  As the bodies erode, wide
channels form where streamlines undergo only minor vertical deflections,
and this explains why the tortuosity eventually decreases with porosity.
We computed lines of best fit using the porosity-tortuosity
models~\eqref{eqn:tortuosityModels} and found the power law to give the
smallest error.  The  black dashed line in
Figure~\ref{fig:Eroding20Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064}
\end{align}
with a root-mean-square error of $5.90 \times 10^{-3}$.  Interestingly,
at the low porosities, the tortuosity initially increases. This increase
occurs because in the absence of erosion (left plot in
Figure~\ref{fig:Eroding20vort}), many of the streamlines, such as those
initialized between bodies 15 \& 18, only perform minor deflections to
pass through the narrow regions, albeit, very slowly.  However, as
erosion starts to open the channels, the streamlines deflect into the
fast regions, such as the region above body 11, and this increases the
amount of vertical deflection, and therefore the tortuosity. While this
increase in tortuosity is interesting, in the next two examples we will
see that the tortuosity does not initially increase.

In Figure~\ref{fig:Eroding20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  The spreading is
computed for seven geometries of different porosities that are formed
during the erosion process.  So that the spreading reaches a statistical
equilibrium, we use the reinsertion algorithm described in
Section~\ref{sec:dispersion} to form sufficiently long trajectories.
For all the reported porosities, the particle dispersion exhibits two
distinct power law regimes.  Initially, the dispersion is ballistic
($\sigma_\lambda \sim t$) since individual fluid particles have not yet
explored enough space to significantly alter their velocity.  However,
once the particles have been subjected to a range of velocities, their
dispersion slows, and we observe super-dispersive ($\sigma_\lambda \sim
t^\alpha$, $\alpha \in (1/2,1)$) behavior over at least one order of
magnitude in time.  Before any erosion takes place, the anomalous
dispersion coefficient is $\alpha = 0.56$.  Then, as the grains begin to
erode, the dispersion rate grows towards the ballistic regime that
occurs in the absence of grains.  The monotonic increase in dispersion
with respect to the porosity is explained by the onset of channels where
many tracers experience less variability in their velocities.

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/20b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20Transport} (a) The tortuosity of an eroding
geometry initialized with 20 grains.  The tortuosity is calculated using
the Eulerian method~\eqref{eqn:tortuosity2} (blue dots) and Lagrangian
method~\eqref{eqn:tortuosity1} (red stars).  The red square corresponds
to the geometry in Figure~\ref{fig:Eroding20tort}(c).  The dash line is
the line of best fit $\widehat{T}(\phi)=\phi^{-p}$ with $p=0.2064$, and
the corresponding root-mean-square error is $5.90 \times 10^{-3}$. (b)
The temporal evolution of $\sigma_\lambda$ as a function of time at
seven different porosities.  The dashed line, which has slope one,
corresponds to ballistic dispersion $\sigma_\lambda \sim t$ that occurs
if no grains are present. At early times, the particles undergo a
ballistic motion, but once they have traversed a few grains, the
spreading becomes super-dispersive with $\sigma_\lambda \sim t^\alpha$,
$\alpha \in (1/2,1)$.  The dashed-dotted lines are lines of best fit
with slopes $\alpha = 1.06$ ($\phi=95.10\%$), $\alpha = 1.07$
($\phi=85.09\%$), $\alpha = 1.06$ ($\phi=75.15\%$), $\alpha = 0.97$
($\phi=65.09\%$), $\alpha = 0.78$ ($\phi=55.10\%$), $\alpha = 0.75$
($\phi=45.08\%$), and $\alpha = 0.56$ ($\phi=37.68\%$).  Values greater
than 1 are a result of using a least-squares fit for the tails of the
particle spreading.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Low Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a second example with 20 eroding bodies, but with a smaller
initial porosity.  In Figure~\ref{fig:ErodingLow20vort}, we plot the
eroding geometry and vorticity at four evenly spaced instances in time.
Initially, the smallest distance between pairs of bodies is $3.29 \times
10^{-4}$, and the smallest distance between the bodies and solid wall is
$4.50 \times 10^{-3}$.  At these distances, a resolution of
approximately $N_\iin = 27,000$ and $N_\out = 18,000$ discretization
points is required to satisfy the $5h$ threshold needed for the
trapezoid rule to achieve machine precision.  By applying the
Barycentric quadrature rule, we stably simulate erosion at the much
coarser resolution of $N_\iin = 256$ and $N_\out = 1024$ discretization
points.

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort1}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort150}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort300}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort450}
\end{center}
\caption{\label{fig:ErodingLow20vort} The erosion of 20 nearly touching
bodies, each discretized with $N_\iin = 256$ points, in a
Hagen-Poiseuille flow. The four snapshots are evenly spaced in time, and
the color is the fluid vorticity. In addition to the channels that
develop between the bodies and the solid walls, erosion leads to two
main channels through the geometry---one in the top half and one near
the middle.}
\end{figure}

To avoid computing streamlines at every time step during the erosion
process, we compute the tortuosity using the Eulerian
method~\eqref{eqn:tortuosity2}.  The initial porosity is $30.67\%$ and
the initial tortuosity is $T = 1.24$.  In
Figure~\ref{fig:ErodingLow20Transport}(a), we plot the tortuosity with
respect to the porosity (blue) and the line of best fit (black) using
the power law 
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.1669}.
\end{align}
This model outperformed the other three models in
equation~\eqref{eqn:tortuosityModels}, and its root-mean-squared error
is $1.13 \times 10^{-2}$.  As grains erode, there is an increase in the
number of streamlines that take a nearly direct path through the
geometry, and this decreases the tortuosity. However, the channelization
effect of erosion results in an increase in the tortuosity since the
length of many of the streamlines increases when they deflect from a
high porosity region (low pressure) to a low porosity region (high
pressure). For this example, we see that the net effect is a decrease in
the tortuosity for all time.

In Figure~\ref{fig:ErodingLow20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  As in the last
example, we analyze the spreading at several different porosities and we
use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories to observe an asymptotic
dispersion rate.  For all the porosities, the dispersion is much closer
to ballistic when compared to the results in
Figure~\ref{fig:Eroding20Transport}.  However, there are still clear
transitions from ballistic dynamics to asymptotic super-dispersive
spreading.  In contrast to the higher porosity initial condition
(Section~\ref{sec:Eroding20}), at early times the erosion results in a
decrease in the dispersion rate. In particular, the first 5\% of the
bodies have eroded, the particle spreading transitions from
$\sigma_\lambda \sim t^{0.92}$ to $\sigma_\lambda \sim t^{0.69}$.  To
explain this behavior, recall that anomalous dispersion is caused by
tracers spending time in both the fast and slow regimes.  Since the
initial configuration has a reasonably uniform velocity (see
Figure~\ref{fig:ErodingLow20vort}), albeit a small one, the dispersion
is nearly ballistic. However, as the geometry erodes, the flow becomes
more intermittent, and this results in an increased anomalous dispersion
rate~\cite{dea-leb-den-tar-bol-dav2013}.  Then, as the bodies continue
to erode, the geometry channelizes, and most tracers are transported
with a large velocity through the channels, again resulting in a nearly
ballistic motion~\cite{sie-ili-pri-riv-gua2019}.

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian20b}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/20b_dense_second_moment_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:ErodingLow20Transport} (a) The tortuosity of an
eroding geometry initialized with 20 grains.  When compared to the last
example, the bodies are initially much closer together, and the porosity
is smaller.  The tortuosity is calculated with the Eulerian
method~\eqref{eqn:tortuosity2} (blue dots).  The dash line is the
fitting line $\widehat{T}(\phi)=\phi^{-p}$ with $p=0.1669$, and the root
mean square deviation is $1.13 \times 10^{-2}$.  (b) The temporal
evolution of $\sigma_\lambda$ as a function of time at eight different
porosities.  The dashed line, which has slope one, corresponds to
ballistic dispersion $\sigma_\lambda \sim t$ that occurs if no grains
are present.  At early times, the particles undergo a ballistic motion,
but once they have traversed a few grains, the spreading becomes
super-dispersive with $\sigma_\lambda \sim t^{\alpha}$, $\alpha \in
(1/2,1)$.  The dashed-dotted lines are lines of best fit with slope
$0.92$ ($\phi=95.00\%$), $0.87$ ($\phi=85.17\%$), $0.87$
($\phi=75.10\%$), $0.91$ ($\phi=65.02\%$), $0.91$ ($\phi=55.08\%$),
$0.72$ ($\phi=45.03\%$), $0.69$ ($\phi=35.05\%$), and $0.92$
($\phi=30.67\%$).}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{100 eroding bodies}
\label{sec:Eroding100}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As a final example, we consider 100 eroding bodies with an initial
porosity near 50\%.  Snapshots of the configurations and vorticity are
in Figure~\ref{fig:Eroding100vort}.  Again, by resolving near-contacts
with the Barycentric quadrature rule, a stable simulation is possible
with only $N_\iin = 256$ discretization points.

\begin{figure}[H]
 \begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =0.9\linewidth]{./figs/100b_50}
\caption{100 bodies, $\phi = 55.54\%$}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =1.1\linewidth]{./figs/100b_100}
\caption{99 bodies, $\phi = 62.98\%$}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =0.9\linewidth]{./figs/100b_150}
\caption{94 bodies, $\phi = 72.22\%$}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =1.1\linewidth]{./figs/100b_200}
\caption{82 bodies, $\phi = 83.67\%$}
\end{subfigure}
\caption{\label{fig:Eroding100vort} The erosion of 100 nearly
touching grains, each discretized with $N_\iin = 256$ points, in a
Hagen-Poiseuille flow. The color is the vorticity of fluid. The four
snapshots are evenly spaced in time, and the color is the fluid
vorticity.  Because of the large number of bodies, erosion creates many
channels connecting the inlet to the outlet.}
\end{figure}

We compute the tortuosity using both the
Lagrangian~\eqref{eqn:tortuosity1} and Eulerian
methods~\eqref{eqn:tortuosity2}. Therefore, we must compute the velocity
at $N_p = 1000$ points along the inlet $x=-1$. These normalized
velocities are plotted in Figure~\ref{fig:Eroding100tort}(a) for the
eroded geometry at porosity $\phi = 62.98\%$
(Figure~\ref{fig:Eroding100tort}(c)).  The initial velocity of the
tracers is qualitatively similar to the 20 body example
(Figure~\ref{fig:Eroding20tort}(a)), except with additional maximum and
minimums because of the additional bodies.  Next, in
Figure~\ref{fig:Eroding100tort}(b), we plot the local tortuosity by
finding the length of the streamlines as they pass from $x=-1$ to $x=1$.
Compared to Figure~\ref{fig:Eroding20tort}(b), the local tortuosity is
much more discontinuous.  These discontinuities can be explained by
examining the trajectories of tracers in
Figure~\ref{fig:Eroding100tort}(c).  Here, there are many instances of
nearby streamlines that are deflected apart from one another as they
tend to a stagnation point in the flow, and this results in trajectories
with significantly different lengths.  At this porosity, one of the
tracers travels 25.5\% farther than it would have if the bodies had been
absent, and the average tracer travelled 12\% farther indicating a
tortuosity of $1.12$.

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local100}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width = \textwidth]{./figs/100b_t100tracer}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100tort} The tortuosity of a porous geometry
that initially had 100 grains and has eroded to a porosity of 62.98\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1, y)$,
normalized by its maximum velocity $u_{max}=3.90 \times 10^{-4}$.  Note
that this maximum velocity is about a order of magnitude smaller than
when compared to the 20 body example in Figure~\ref{fig:Eroding20tort}.
(b) The local tortuosity $\tau(y)$ on the cross section $x = -1$. The
spatially averaged tortuosity through the channel is $1.12$. (c) The
trajectories of 200 tracers initialized at $x = -1$.  The
discontinuities in the local tortuosity correspond to nearby streamlines
whose lengths are significantly different because they pass opposite
sides of an eroding body. Compared to Figure~\ref{fig:Eroding20tort},
this example has more small bodies, and this results in many more
discontinuities.}
\end{figure}

In Figure~\ref{fig:Eroding100Transport}(a), we plot the tortuosity as a
function of the porosity.  The initial geometry has a porosity of $\phi
= 50.09\%$ and the tortuosity is $T = 1.20$.  The tortuosity is computed
with both the length of the streamlines~\eqref{eqn:tortuosity1} (red
stars) and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in Figure~\ref{fig:Eroding100tort}(c).
Again, the two tortuosity formulas give similar results.  For this
geometry, the tortuosity decreases monotonically at almost all
porosities.  However, the tortuosity undergoes a sudden increase near
the end of the simulation, and we have observed this behavior in other
examples.  The increase is caused by a single small body near the middle
of the channel being completely eroded.  While this results in
straighter streamlines, therefore reducing $\lambda$, the horizontal
flow, $u_1(y)$ increases since there is no longer a no-slip boundary.
We also computed the lines of best fit using the porosity-tortuosity
models~\eqref{eqn:tortuosityModels}.  The black dashed line in
Figure~\ref{fig:Eroding100Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2459}
\end{align}
with a root-mean-square error of $5.50 \times 10^{-3}$.  We note a
slightly better root-mean-square error of $5.20 \times 10^{-3}$ is
possible with the model
\begin{align}
  \widehat{T}(\phi) = 1 - 0.2631 \ln(\phi).
\end{align}

In Figure~\ref{fig:Eroding100Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$ at six different
porosities. Again, we initially observe ballistic motion (black dashed
line), and then super-dispersion.  Similar to the example in
Figure~\ref{fig:ErodingLow20Transport}(b), the anomalous dispersion rate
is not growing with the porosity.  Therefore, it appears that the
dispersion rate in an eroding geometry depends not only the porosity,
but also the location and shape of the bodies.  Finally, at the highest
porosity, anomalous dispersion is only observed briefly in the time
interval $(0.5,1)$, and then transitions back to a ballistic regime.
Since the bodies are so small at this high porosity, after reinsertion,
the streamline is not significantly deflected by any of the bodies, and
this results in a ballistic regime.

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/100b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100Transport} (a) The tortuosity of an
eroding geometry initialized with 100 grains.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue dots)
and Lagrangian method~\eqref{eqn:tortuosity1} (red stars).  The dash
line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$ with $p=0.2459$,
and the root mean square deviation is $5.50 \times 10^{-3}$.  We note
that the model $\widehat{T}(\phi) = 1 - p \ln (\phi)$, with $p=0.2631$,
has a comparable root-mean-square error of $5.20 \times 10^{-3}$.  (b)
The temporal evolution of $\sigma_\lambda$ as a function of time at six
different porosities.  The dashed line, which has slope one, corresponds
to ballistic dispersion $\sigma_\lambda \sim t$ that occurs if no bodies
are present.  At early times, the particles undergo a ballistic motion,
but once they have traversed a few grains, the spreading becomes
super-dispersive with $\sigma_\lambda \sim t^{\alpha}$, $\alpha \in
(1/2,1)$.  The dashed-dotted lines are lines of best fit with slope
$1.11$ ($\phi=95.12\%$), $0.68$ ($\phi=85.07\%$), $0.73$
($\phi=75.14\%$), $0.59$ ($\phi=65.03\%$), $0.61$ ($\phi=55.02\%$), and
$0.70$ ($\phi=50.09\%$).}
\end{figure}

Finally, we investigate the effects of erosion on pore sizes.  The
distribution of the pore sizes can be directly related to the
distribution of the velocity, and thus effects the
tortuosity~\cite{den-ica-hid2018} and anomalous
dispersion~\cite{dea-qua-bir-jua2018}. In addition, the pore sizes are
required in network models~\cite{bry-mel-cad1993, bry-kin-mel1993}.  As
described in Section~\ref{sec:throats}, we use a Delaunay triangulation
to define neighboring eroding bodies, and we compute the pore size by
finding the closest distance between all neighboring bodies.  Instead of
computing a Delanuay configuration at each time step, which would result
in new definitions for the pores at each time step, we only compute a
new Delaunay triangulation when a grain completely erodes.  Once all
pore sizes are computed, we analyze their distribution as a function of
the porosity.

HERE
In Figure~\ref{fig:Eroding100gap_hist}, we plot
histograms of the pore pore sizes at six porosities during the erosion
process. These types of distributions have been described with Weibull
distributions~\cite{ioa-cha1993}, and we superimpose the Weibull
distribution with the same first two moments as the data.  The
parameters of the distribution, $(k,\lambda)$, are included in the
caption of Figure~\ref{fig:Eroding100gap_hist}.  In
Figure~\ref{fig:Eroding100gap_mean_var}, we plot the mean and variance
of the pore sizes as a function of the porosity.  Interestingly, for
porosities less than $\phi = 85\%$, the mean pore size grows linearly
and the variance remains nearly flat. Since a channelized geometry has
large variance, this indicates that channelization is less prevalent at
low porosities.

\begin{figure}[H]
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_1}
\caption{100 bodies, 258 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_46}
\caption{100 bodies, 258 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_112}
\caption{97 bodies, 249 pores}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_164}
\caption{92 bodies, 234 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_207}
\caption{75 bodies, 189 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_246}
\caption{34 bodies, 74 pores}
\end{subfigure}
\caption{\label{fig:Eroding100gap_hist} The distribution of the pore
sizes of 100 eroding bodies at six porosities. The black curves are
Weibull distributions whose first and second moments agree with the
data. The porosities and Weibull distribution parameters $(k,\lambda)$
at each of the porosities are: (a) $\phi = 50.09\%$ and
$(k,\lambda)=(1.4650,0.0605)$; (b) $\phi = 55.02\%$ and
$(k,\lambda)=(1.9485,0.0737)$; (c) $\phi = 65.03\%$ and
$(k,\lambda)=(2.5682,0.0965)$; (d) $\phi = 75.14\%$ and
$(k,\lambda)=(2.7771,0.1255)$; (e) $\phi = 85.07\%$ and
$(k,\lambda)=(2.6235,0.1713)$; (f) $\phi = 95.12\%$ and
$(k,\lambda)=(1.9840, 0.3194)$.}
\end{figure}

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_mean}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/gap_variance}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap_mean_var} The effect of erosion on the
mean (a) and the variance (b) of the pore sizes. The geometry initially
contains 100 eroding bodies.  The distributions of the pore sizes in
Figure~\ref{fig:Eroding100gap_hist} are indicated by the red stars.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As a continuation of our previous work~\cite{qua-moo2018}, we have
simulated dense suspensions and characterized transport in viscous
erosion of porous media. A BIE formulation is used to solve the Stokes
equations and to compute the shear stress. However, to simulate
nearly-touching grains and to consider transport, it is necessary to
integrate near-singular functions. This is achieved by extending a
Barycentric quadrature rule~\cite{bar-wu-vee2015} to accurately compute
the erosion rate and vorticity. By using this quadrature rule, we are
able to perform stable simulations of erosion with $N =
\mathcal{O}(100)$ discretization points, while the trapezoid rule would
require $\mathcal{O}(10^5)$ discretization points.

The transport is characterized in terms of tortuosity and anomalous
dispersion. The local tortuosity agrees qualitatively with other
works~\cite{mat-kha-koz2008}. However, the tortuosity of the eroded
geometries cannot be completely described in terms of the porosity. In
particular, we observe that for certain porous media, the tortuosity can
increase when the porosity increases under erosion. We also observe
super-dispersive spreading, and the rate of dispersion significantly
depends not only on the porosity, but also the number of eroding bodies.

To further our understanding of erosion, we are examining other bulk and
statistical properties of an eroding porous media. In this work, we
provide preliminary results for the pore throat sizes which affect the
anomalous dispersion rate~\cite{dea-qua-bir-jua2018}.  At a later date,
we will report results on the  development of anisotropic effects and
the distributions of grain sizes, shapes, and opening angles.

As a long term goal, we would like to include the effect of inertia and
other transport models. Including inertia requires an integral equation
formulation of the Navier-Stokes equations, which is an active area of
research with a few promising algorithms recently
proposed~\cite{gray2019boundary, kli-ask-kro2019}.  Regarding other
transport models, this would involve a diffusive term to consider the
transport of heat or a contaminant. Forming high-fidelity simulations of
such an advection-diffusion equation can be accomplished by using time
splitting methods and recent work on heat solvers in complex
geometries~\cite{fry-kro-tor2019}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{\bf Acknowledgments} BQ and NM were supported by Florida
State University startup funds and Simons Foundation Mathematics and
Physical Sciences-Collaboration Grants for Mathematicians.

\bibliographystyle{plainnat} 
\biboptions{sort&compress}
\bibliography{refs}

%old stuff%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{boundary integral equation formulation} 
%\label{sec:bies}
%add an introduction why we do not use the double-layer potentials of stokes equations but the double-layer potentials of laplace equations to solve stokes equation when we use barycentric approach (or complex variable approach).
%
%\subsubsection{laplace's equations from $\rdb^2$ to $\cdb$ }
%
%let $\omega$ be the domain of fluid. the boundary of $\omega$ is $\gamma$ and the outward normal vector on $\gamma$ is ${\bf n_y}=(n_1,n_2)$.
%
%consider a laplace's equations with dirichlet boundary condition
%\begin{equation}\label{laplace}
%  \begin{cases}
%    \delta {\bf u}=0, & \text{in}\,\,\, \omega \in \rdb^2;\\
%    {\bf u}={\bf f}, & \text{on}\,\,\, \gamma.
%  \end{cases}
%\end{equation}
%
%Let ${\bf x}=(x_1,x_2), {\bf y}=(y_1,y_2)$. Then we have ${\bf r}={\bf x} - {\bf y}$ and $\rho=|\bf r|$.
%The double-layer potential which solves \eqref{Laplace} is 
%\begin{align*}
%\D[\sigma]({\bf x})&=\frac{1}{2\pi}\displaystyle\int_{\Gamma} \frac{{\bf r} \cdot {\bf n_y}}{\rho^2}\sigma({\bf y})ds_y,
%%\frac{1}{2\pi}\displaystyle\int_{\Gamma}\frac{\partial}{\partial {\bf n_y}} \log |{\bf x}-{\bf y}| \sigma({\bf y})ds_y,\\ \\
%%&=\frac{-1}{2\pi}\int_{\Gamma}\frac{{\bf r}\cdot{\bf n_y}}{\rho^2} \sigma({\bf y})ds_y\\ \\
%%&=\frac{-1}{2\pi}\int_{\Gamma}\frac{n_1(x_1-y_1)+n_2(x_2-y_2)}{(x_1-y_1)^2+(x_2-y_2)^2} \sigma({\bf y})ds_y 
%\numberthis\label{DLP:real}
%\end{align*}
%where $\sigma$ is the density function on the boundary $\Gamma$.
%
%
%Consider a Cauchy integral 
%\begin{equation}\label{CI}
%v({x})=\frac{1}{2\pi i}\int_{\Gamma}\frac{\sigma({ y})}{{ y}-{ x}} d{ y},\,\,\,\,\,\,\,\,\, \forall { x} \in \Cdb\setminus \Gamma.
%\end{equation}
%where ${ x}=x_1+ix_2$, $ { y}=y_1+iy_2$, ${ n_y}=n_1+i n_2$ and $d{ y}=i {n_y}ds_y$. It can be shown that 
%the real part of the Cauchy integral in \eqref{CI} is equivalent to the double-layer potential in \eqref{DLP:real}. 
%
%
%\subsubsection{From Stokes DLP to Laplace DLPs}
%
%Let ${\bf x}=(x_1,x_2), {\bf y}=(y_1,y_2)$. Then we have ${\bf r}={\bf x} - {\bf y}$ and $\rho=|\bf r|$.
%
%
%The Stokes double-layer potential  is
%\begin{equation}\label{StokesDLP1}
%\D[\boldsymbol\sigma]({\bf x}):=\frac{1}{\pi}\int_{\Gamma}\left(\frac{{\bf r} \cdot {\bf n_y}}{\rho^2}\frac{{\bf r} \otimes {\bf r}}{\rho^2}\right) {\boldsymbol\sigma({\bf y})} d s_y.
%\end{equation}
%We could rewrite the Stokes DLP in terms of the Laplace DLP as
%\begin{align*}\label{StokesDLP2}
%\D[\boldsymbol\sigma]({\bf x})=\frac{1}{2\pi}\int_{\Gamma}&\frac{{\bf n_y}}{\rho^2}({\bf r} \cdot {\boldsymbol \sigma})ds_y+\frac{1}{2\pi}\nabla\int_{\Gamma}\frac{{\bf r} \cdot {\bf n_y}}{\rho^2}({\bf y} \cdot{\boldsymbol \sigma})ds_y\\
%&-\frac{1}{2\pi}x_1 \nabla\int_{\Gamma}\frac{{\bf r} \cdot {\bf n_y}}{\rho^2}\sigma_1ds_y-\frac{1}{2\pi}x_2\nabla\int_{\Gamma}\frac{{\bf r} \cdot {\bf n_y}}{\rho^2}\sigma_2ds_y. \numberthis
%\end{align*}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Computing the shear stress}
%\label{sec:shearStressLP}
%
%To calculate the shear stress, we need to know the gradient of double-layer potential in \eqref{StokesDLP2} 
%\begin{align*}\label{SSLP}
%\nabla\D[\boldsymbol\sigma]({\bf x})=\frac{1}{2\pi}\nabla\int_{\Gamma}&\frac{{\bf n_y}}{\rho^2}({\bf r} \cdot {\boldsymbol \sigma})ds_y+\frac{1}{2\pi}\nabla^2\int_{\Gamma}\frac{{\bf r} \cdot {\bf n_y}}{\rho^2}({\bf y} \cdot{\boldsymbol \sigma})ds_y\\
%&-\frac{1}{2\pi}\nabla\left(x_1 \nabla\int_{\Gamma}\frac{{\bf r} \cdot {\bf n_y}}{\rho^2}\sigma_1ds_y\right)-\frac{1}{2\pi}\nabla\left(x_2\nabla\int_{\Gamma}\frac{{\bf r} \cdot {\bf n_y}}{\rho^2}\sigma_2ds_y\right). \numberthis
%\end{align*}
%
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Boundary evolution in the {\thL} framework} 
%\label{sec:thetaL}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Solving the BIE}
%\label{sec:BIE}
%\subsubsection{Interior Case}
%Consider a Cauchy integral 
%\begin{align}\label{CIFI}
%\frac{1}{2\pi i}\int_{\Gamma}\frac{v^-({ y})}{{ y}-{ x}} d{ y}
%&=\begin{cases}
%\,v(x), &{ x} \in \Omega,\\ 
%\,0,  &{ x} \in \Cdb\setminus \bar{\Omega}.
%\end{cases}
%\end{align}
%By the identity of Cauchy Integral Formula $\displaystyle\frac{1}{2 \pi i}\int_{\Gamma}\frac{1}{{ y}-{ x}} d{ y}=1,$ we can rewrtie \eqref{CIFI} and get 
%\begin{equation}
%\frac{1}{2\pi i}\int_{\Gamma}\frac{v^-({ y})-v(x)}{{ y}-{ x}} d{ y}=0,\,\,\,\,\,\,\,\,\, \forall { x} \in \Omega.
%\end{equation}
%\begin{equation}\label{dNCI}
%v'({x})=\frac{1}{2\pi i}\int_{\Gamma}\frac{v^-({ y})}{(y- x)^2} d{ y},\,\,\,\,\,\,\,\,\, \forall { x} \in \Omega.
%\end{equation}
%By the identity of Cauchy Integral Formula $\displaystyle\frac{1}{2 \pi i}\int_{\Gamma}\frac{1}{( y- x)^n} d{ y}=0$ for integer n $\neq 1$, we can rewrtie \eqref{dNCI} and get 
%\begin{equation}
%\frac{1}{2\pi i}\int_{\Gamma}\frac{v^-({ y})-v(x)-(y-x)v'(x)}{(y-x)^2} d{ y}=0,\,\,\,\,\,\,\,\,\, \forall { x} \in \Omega.
%\end{equation}
%
%\subsubsection{Exterior case}
%Consider a Cauchy integral formula for the exterior domain
%\begin{align}\label{CIFE}
%\frac{1}{2\pi i}\int_{\Gamma}\frac{v^+({ y})}{{ y}-{ x}} d{ y}
%&=\begin{cases}
%\,v_{\infty}, &{ x} \in \Omega,\\ 
%\,-v(x)+v_{\infty},  &{ x} \in \Omega^c,
%\end{cases}
%\end{align}
%where $v_{\infty}=\displaystyle\lim_{x \to \infty}v(x)$.
%\begin{equation}
%\frac{1}{x-a}=\frac{-1}{2 \pi i} \int_{\Gamma}\frac{(y-a)^{-1}}{y-x} d{ y},\,\,\,\,\,\,\,\,\, \forall { x} \in \Omega^c.
%\end{equation}
%\begin{equation}
%\frac{1}{2\pi i}\int_{\Gamma}\frac{v^+({ y})-(y-a)^{-1}(x-a)v(x)}{y-x} d{ y}=0,\,\,\,\,\,\,\,\,\, \forall { x} \in \Omega^c.
%\end{equation}
%\begin{equation}
%\frac{1}{2\pi i}\int_{\Gamma}\frac{v^+({ y})-v(x)-(y-a)^{-1}(x-a)(y-x)v'(x)}{(y-x)^2} d{ y}=0,\,\,\,\,\,\,\,\,\, \forall { x} \in \Omega^c.
%\end{equation}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Shear stress}
%\label{sec:shearStress}
%\subsubsection{Interior Case}
%\begin{equation}\label{ddNCI}
%v''({x})=\frac{1}{2\pi i}\int_{\Gamma}\frac{2v^-({ y})}{(y- x)^3} d{ y},\,\,\,\,\,\,\,\,\, \forall { x} \in \Omega.
%\end{equation}
%By the identity of Cauchy Integral Formula $\displaystyle\frac{1}{2 \pi i}\int_{\Gamma}\frac{1}{( y- x)^n} d{ y}=0$ for integer n $\neq 1$, we can rewrtie \eqref{ddNCI} and get 
%\begin{equation}
%\frac{1}{2\pi i}\int_{\Gamma}\frac{2v^-({ y})-2v(x)-2(y-x)v'(x)-(y-x)^2v''(x)}{(y-x)^3} d{ y}=0,\,\,\,\,\,\,\,\,\, \forall { x} \in \Omega.
%\end{equation}
%\subsubsection{Exterior case}
%\begin{equation}
%\frac{1}{2\pi i}\int_{\Gamma}\frac{2v^+({ y})-2v(x)-2(y-x)v'(x)-(y-a)^{-1}(x-a)(y-x)^2v''(x)}{(y-x)^3} d{ y}=0,\,\,\,\,\,\,\,\,\, \forall { x} \in \Omega^c.
%\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% TIME-STEPPING %%
%\subsection{Time-stepping with the {\thL} method} 
%\label{sec:timeStepping}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Post processing quantities of interest}
%\label{s:qoi}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Vorticity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Pressure}
%\label{sec:pressure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Drag}
%\label{sec:drag}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Near-singular integration}
%\label{sec:NSI}
%
%
%{\color{red}
%Here we derive the equations~\eqref{eqn:pderiv1}-\eqref{eqn:pderiv2} 
%and the other two are similar.
%To derive equations~\eqref{eqn:pderiv1} and \eqref{eqn:pderiv2}, 
%we need to apply the fact that if a real function $u$ is equivalent to 
%the real part of a complex varaible function $f$, 
%then we have $\nabla u=(\Re(f'),-\Im(f'))$.
%Let $f_1(x)= v[\tau_1](x)$, $f_2(x)=v'[\yy\cdot\eeta](x)$, $f_3(x)=v'[\eta_1](x)$,
%and $f_4(x)= v'[\eta_2](x)$. Then we can rewrite equation~\eqref{eqn:cauchy1}
%\begin{align*}
%u_1(\xx)&= u_{11}(\xx) +  u_{12}(\xx) +u_{13}(\xx)+u_{14}(\xx) \\
%&=\Re (f_1(x)) + \Re (f_2(x))-x_1\Re (f_3(x)) - x_2\Re (f_4(x)).
%\end{align*}
%To get the gradient of $u_1(x)$, we have
%\begin{align*}
%\nabla u_1(x)&= \nabla u_{11}(x)+\nabla  u_{12}(x) 
%+\nabla u_{13}(x)+\nabla u_{14}(x) \\ \\
%&=\begin{bmatrix}
%\Re(f_1'(x))\\ \\
%-\Im(f_1'(x))
%\end{bmatrix}+
%\begin{bmatrix}
%\Re(f_2'(x))\\ \\
%-\Im(f_2'(x))
%\end{bmatrix}
%+
%\begin{bmatrix}
%-\Re(f_3(x))-x_1\Re(f_3'(x))\\ \\
%x_1\Im(f_3'(x))
%\end{bmatrix}
%+
%\begin{bmatrix}
%-x_2\Re(f_4'(x))\\ \\
%-\Re(f_4(x))+x_2\Im(f_4'(x))
%\end{bmatrix}\\ \\
%&=
%\begin{bmatrix}
%\Re (v'[\tau_1](x)) + 
%    \Re (v''[\yy\cdot\eeta](x)) - \Re (v'[\eta_1](x)) - 
%    x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x))\\ \\
%- \Im (v'[\tau_1](x)) - 
%    \Im (v''[\yy\cdot\eeta](x)) + x_1\Im (v''[\eta_1](x)) - 
%    \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x))
%\end{bmatrix}.
%\end{align*}
%
%{\color{red}
%\begin{alignat}{3}
%1 &= \frac{1}{2\pi i}\int_{\Gamma} 
%    \frac{1}{y-x} dy, &&x \in \Omega, \label{eqn:cauchyId1}\\
%\frac{1}{x-a} &= \frac{-1}{2\pi i}\int_{\Gamma} 
%    \frac{(y-a)^{-1}}{y-x} dy,  \quad &&x \in \Omega^c\label{eqn:cauchyId2}. 
%\end{alignat}
%and the contour integral 
%\begin{align}\label{eqn:cauchyId3}
%0=\oint_{\Gamma} \frac{1}{(y-x)^n}dy, \quad n \neq 1.
%\end{align}
%Then we apply equations~\eqref{eqn:cauchyId1} and~\eqref{eqn:cauchyId3} on
%the second derivative of Cauchy integral in \eqref{eqn:cauchy} and get
%\begin{align*}
%0 &= \frac{1}{2\pi i} \int_{\Gamma}
%    \frac{2v^-(y)}{(y - x)^3} \, dy -v''(x)\\
%&= \frac{1}{2\pi i} \int_{\Gamma}
%    \frac{2v^-(y)}{(y - x)^3} \, dy 
%-v''(x)\frac{1}{2\pi i}\int_{\Gamma} 
%    \frac{1}{y-x} dy\\
%&= \frac{1}{2\pi i} \int_{\Gamma}
%    \frac{2v^-(y)}{(y - x)^3} \, dy 
%-\frac{1}{2\pi i}\int_{\Gamma} 
%    \frac{(y-x)^2v''(x)}{(y-x)^3} dy
%-\frac{1}{2\pi i}\int_{\Gamma} \frac{2v(x)}{(y-x)^3}dy
%-\frac{1}{2\pi i}\int_{\Gamma} \frac{2(y-x)v'(x)}{(y-x)^3}dy\\
%&= \frac{1}{2\pi i}\int_{\Gamma} 
%    \frac{2v^-({ y})-2v(x)-2(y-x)v'(x)-(y-x)^2v''(x)}{(y-x)^3} dy.
%\end{align*}
%}
%\todo[inline]{S-H: How do we get these identities?}
%\begin{alignat}{3}
%  0 &= \frac{1}{2\pi i}\int_{\Gamma} 
%    \frac{2v^-({ y})-2v(x)-2(y-x)v'(x)-(y-x)^2v''(x)}{(y-x)^3} dy, &&x \in \Omega, \\
%  0 &= \frac{1}{2\pi i}\int_{\Gamma}
%    \frac{2v^+(y)-2v(x)-2(y-x)v'(x)-(y-a)^{-1}(x-a)(y-x)^2 v''(x)}
%    {(y-x)^3} dy, \quad &&x \in \Omega^c.  \label{eqn:Identity}
%\end{alignat}
%{\color{red}
%and the quadrature formulations are
%\begin{alignat}{3}
%  v''(x) &\approx
%    \frac{2\sum\limits_{j=1}^N \frac{v_j^- -v(x)-(y_j-x)v'(x)}{(y_j-x)^3}w_j}
%    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, &&x \in \Omega, \\
%  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
%    \frac{v_j^+ -v(x)-(y_j-x)v'(x)}{(y_j-x)^3}}{\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}
%    , \quad &&x \in \Omega^c.  \label{eqn:SecondBary}
%\end{alignat}
%where $a$ is a point inside the $\Omega$ in equations~\eqref{eqn:Identity} and \eqref{eqn:SecondBary}.
%
%}
%
%{\color{red}
%By using Barycentric quadrature rule to approach integrations, we get a reasonable(?) double-layer potential and quantities of interest at target points but it requires $\mathcal{O}(N^2)$ operations per target point which is computational expensive if we impose it on all the target points. In order to accelerate the process and keep the accuracy, it is a compromise that we use Barycentric quadrature rule to approach the nearly singular integratons but apply trapezoid rule when the target point is far from source points. 
%
%
%To numerically compute the completed double-layer potential, we discretize each of the $M$ bodies with  $N_\iin$ points, and the outer wall with $N_\out$ points. The following is the four-step fast summation methods.
%}
%\begin{itemize}
%  \item Apply trapezoid rule as a first pass
%
%{\color{red} We use the Fast Multipole Method (FMM)~\cite{gre-rok1987, gre-gre-may1992} with trapezoid rule to calculate the double-layer potential.
%}
%  \item Efficiently find points that are too close for trapezoid
%
%{\color{red}
% Let $L_w$ be the length of the outer wall and $L_{b}$ be the length of the boundary of body where $b=1, \cdots, M$. Here are the criteria to see if two points are too close for trapezoid. For criterion of outer wall to body,  let $d$ be the minimum distance between outer wall and a point $x$ on the body $b$, we say $x$ is too close to outer wall if
%\begin{equation}\label{eqn:Cw2b}
%d< \alpha_1\frac{L_w}{N_\out}
%\end{equation}
%where $\alpha_1$ is a constant. 
%For criterion between two bodies, let $d$ be the distance between the centroid $a$ of the body $b_1$ and a point $x$ on the body $b_2$, we say they are too close if
%\begin{equation}\label{eqn:Cb2b}
%d<\alpha_2\left(\frac{L_{b_1}}{N_\iin}+\frac{L_{b_2}}{N_\iin}\right)
%\end{equation}
%where $\alpha_2$ is a constant. 
%
%}
%\end{itemize}
%{\color{red}  (Discussion of the choices of $\alpha_1$ and $\alpha_2$ with respect to accuracy and  efficiency)
%
%To find the suitable values of $\alpha_1$ and $\alpha_2$ in~\eqref{eqn:Cw2b} and \eqref{eqn:Cb2b}, we set $\alpha_1=\alpha_2=1, 2, 3$, and $4$ to test the criteria on a 20-body problem with $N_\out = 1024$ and $N_\iin = 128, 256$, and $512$. To check the accuracy of the simulations, we use the solution calculated by \todo[inline]{what method} to \todo[inline]{...}
%To see how efficient our code is, we record the cpu time for each step.
%
%}

%{\color{red} 
%By using Barycentric quadrature rule, we are able to capture the details
%in the fluid when wall and bodies or body and bodies are close to each
%other.  The first nearly touching bodies case is 20 bodies in a
%Hagen-Poiseuille flow.  The initial velocity of flow from the left side
%of $\Omega$ is $$\UU(\xx)=U\begin{bmatrix} 1-y^2\\ 0\end{bmatrix},$$
%where $U$ is chosen such that the pressure drop between  $x=-2$ and
%$x=2$ remains fixed at 8.  With this choice, the velocity in the absent
%of grains has flow rate $U=1$.  The number of discretization points is
%$N_\iin = 256$.
%
%The snapshots of the process of 20 nearly touching bodies eroding are in
%Fig.~\ref{fig:Eroding20vort}.  We initially place some bodies to nearly
%contact not only one another but also the outer wall. For example,
%bodies 1, 6, 13, and 15 are very close to outer wall: the distance
%between body 1 and outer wall is $1.3025h$, the distance between body 6
%and outer wall is $2.9275h$, the distance between body 13 and outer wall
%is $2.8447h$, the distance between body 15 and outer wall is $1.3352h$
%where $h=L_w/N_\out$ is the arclength spacing correponding to the outer
%wall.  In~\cite{qua-moo2018} we reported the quick expanding space at
%the narrow channel between neighboring bodies, the developing flat faces
%at the sites of near contact, and bodies vanishing. In this nearly
%touching bodies structure, we also observe these similar phenomena but
%not totally the same.  For the speed of expanding space at the narrow
%channel, we see some narrow channels do expand faster than other wider
%channels.  For example, the spacing between bodies 14 and 18 is smaller
%than the spacing between bodies 7 and 14 initially and the former one is
%comparable to the latter one by the second frame. But there are some
%small initial spacings that do not expand that much at all by the fourth
%frame, for example between bodies 1 \& 6, 3 \& 9.  We could find the
%explanation from the vorticity of fluid in Fig.~\ref{fig:Eroding20vort}.
%Since the eroding velocity on the body is proportional to the absolute
%value of vorticity in our erosion model,  then when the  absolute value
%of vorticity is small there is not much erosion on the body.  In all
%four frames of Fig.~\ref{fig:Eroding20vort}, the vorticity is relatively
%small between bodies 1 \& 6 and 3 \& 9.
%
%Besides the channels between two bodies, there is a single channel goes
%across the domain horizontally and the flow has been concentrated in
%this channel.  In the second frame of Fig.~\ref{fig:Eroding20vort}, this
%channel is between bodies 1 \& 20, 1 \& 9, 4 \& 3, 8 \& 11, 8 \& 2, 6 \&
%12, and 6 \& 19.  We can get a better understanding of this channel from
%the trajectories of tracers in Fig.~\ref{fig:Eroding20tracer}.  In the
%top three frames of  Fig.~\ref{fig:Eroding20tracer}, the flow (or
%tracer) first comes from three small channels between outer wall \& body
%10, bodies 7 \& 14, and bodies 14 \& 18, then the three branches become
%one and go into the main channel between bodies 1 \& 9.  From the first
%frame to the forth frame of Fig.~\ref{fig:Eroding20vort}, this channel
%is expanding faster than other channel and bodies 4 \& 12 has vanished
%and body 19 is almost gone in the forth frame.  The reason of fast
%expansion of this channel is the high vorticity.  In the forth frame of
%Fig.~\ref{fig:Eroding20vort}, we could see the vorticity is relatively
%high on the side of bodies which form this channel so the bodies are
%eroded fast on the side of this channel or have vanished completely.
%
%Furthermore, how long the body lasts for depends on not only the size of
%the body but also the location of the body.  For example, the first
%three vanishing bodies are body 12 (as t=0.0256), 4 (as t=0.0278), then
%17 (as t=0.0287).  Since body 4 is located in the main flow channel so
%the body 4 is vanished earlier than body 17 although the initial size of
%body 4 is larger than the size of body 17.  This channel will keep
%expanding and the bodies around this channel will vanish earlier than
%others.  After body 17, the next 8 vanished bodies with respect to time
%are body 19 (as t=0.0312),  body 7 (as t=0.0325), body 20 (as t=0.0329),
%body 8 (as t=0.0349), body 10 (as t=0.0353), body 3 (as t=0.0356), body
%14 (as t=0.0363), then body 9 (as t=0.0376).  All of these eight bodies
%were around this channel before they are gone.
%
%For the flat faces at the sites of near contact, since we put the body
%be close to one another so we are able to see all the channels between
%two close bodies have flat faces.  Besides the channels between bodies,
%we also observe the flat faces on the bodies which are close to the
%outer wall. For example, we see the flat faces on the top of bodies 1 \&
%6 and the bottom of bodies 13 \& 15.  
%
%Since we use a Barycentric quadrature rule to approximate the velocity
%of tracers and apply the fourth order Runge Kutta method to find the new
%locations of tracers in each step, the tracers can be very close to but
%not penetrate the bodies.  Those body-closed tracers are moving
%perfectly so we are able to observe the details of the interaction of
%bodies and fluid.  Although the direction of flow is from left to right,
%the tracers will not only travel horizontally but also vertically.  In
%the last three frames of Fig.~\ref{fig:Eroding20tracer}, some tracers
%climb up from right to left around the body.  There are zoom windows of
%the climbing tracers trajectories in Fig.~\ref{fig:Eroding20zoom} and
%they are taken from the last three frames of
%Fig.~\ref{fig:Eroding20tracer} with more tracers.  In
%Fig.~\ref{fig:Eroding20zoom} the tracers around the center body first go
%from the channel below the body, climb up to the left, then go to the
%right and join another tracers in the channel above the body. 
%
%In Fig.~\ref{fig:Eroding20area} and Fig.~\ref{fig:Eroding20flowrate} we
%measure the area occupied by bodies and the flow rate $U$ quantifily
%during the eroding. In the beginning, 62.32\% of area of $[-1,1] \times
%[-1,1]$ is occupied by 20 bodies. Then the eroding speed is increasing
%when more and more bodies are vanished in the fluid. 
%
%The flow rate increases exponentially with growth rate close to
%$\ln10^4$. 
%
%
%In Fig.~\ref{fig:Eroding20tort} we use tortuosity to analysis the effect
%of the geometrical structure on fluid velocity field. In
%Fig.~\ref{fig:Eroding20tort}(a) the graph represents the ratio of the
%x-component velocity u(-1, y) to its maximum velocity $u_{max}=6.2281
%\times 10^{-4}$ at equal spreading tracers on the cross section (x=-1
%and y in [-1,1]). The velocity is continuous along the cross section and
%acting as several Poiseuille flows. This velocity graph is different
%from the one in~\cite{matyka2008tortuosity} since we choose our cross
%section to be the beginning of the flow before the fluid touches any
%body. In Fig.~\ref{fig:Eroding20tort}(b) we record the local tortuousity
%$\tau$ of the tracers at the cross section. This graph is not continuous
%and the jumps in $\tau$ mean that the initial neighbor tracers have been
%seperated by the bodies. 
%
%In Fig.~\ref{fig:Eroding20tort}(c)
%In Fig.~\ref{fig:Eroding20tort}(d) 
%}
%
%{\color{red}
%To study the interaction of fluid and fixed bodies in the porous media,
%some parameters can be used to describle the interation quantitively,
%for example the porosity, the permeability, and the tortuosity. 
%The relation of porosity and tortuosity has been studied in~\cite{matyka2008tortuosity}.
%The definition of porosity is the ratio 
%$$\phi=\frac{A_v}{A_T}$$
%where $A_v$ is the area of void-space and $A_T$ is the area of the calculating area.}
%{\color{red} The tortuosity represents the weighted average travel distance of a tracer in a porous media. 
%The definition of tortuosity is 
%$$T=\frac{1}{d}\frac{\displaystyle\int_{S}u_x(y) \lambda(y)dy}{\displaystyle\int_{S}u_x(y)dy}$$
%where $S$ is an arbitrary cross section which is perpendicular to the x axis,
%$\lambda(y)$ is the length of streamline at y on $S$, 
%$u_x(y)$ is the component of velocity at y which is normal to $S$, and 
%$d$ is the distance from $S$ to the end cross section.
%It depends on not only the geometric structure but also the velocity of the fluid 
%which is different from the porosity since porosity only depends on the geometric structure.}
%{\color{red}Beisdes, the local tortuosity~\cite{matyka2008tortuosity} has been used to analysis 
%the local behavior of steamline around a body. 
%The definition of local tortuosity of the tracers at $y$ on $S$ is
%$$\tau(y)=\frac{\lambda(y)}{d}.$$ 
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Cauchy Integral Representation of the Pressure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%The pressure of the double-layer potential is
%\begin{align}
%  p(\xx) = -\frac{1}{\pi} \int_{\bd\Omega} \frac{1}{\rho^2}
%    \left(1 - 2 \frac{\rr \otimes \rr}{\rho^2} \right) \nn
%    \cdot \eeta(\yy) \, ds_\yy.
%\end{align}
%
%
%
%{\color{red}
%We
%measure the area occupied by bodies 
%during the eroding and record the flow rate $U$ quantifily in log scale
%during the eroding in Fig.~\ref{fig:Eroding20flowrate}. 
%Since we simulate the erosion by applying fixed pressure drop 
%the flow rate $U$ is small when there are many bodies in the fluid and 
%when the area fraction is getting smaller $U$ is getting larger.
%In Fig.~\ref{fig:Eroding20flowrate}, 62.32\% of area of $[-1,1] \times
%[-1,1]$ is occupied by 20 bodies initially and the flow rate $U$ starts from $9.86\times 10^{-5}$. 
%Then the eroding speed is increasing
%when more and more bodies are vanished in the fluid and 
%the maximum eroding speed happens when the area fraction equals 7.57\% 
%which is occupied by 8 bodies.
%When all the bodies have vanished completely the flow rate $U$ is equal to 1.
%In Fig.~\ref{fig:Eroding20flowrate}(b), the flow rate $U$ increases exponentially 
%with growth rate close to $\ln10^4$. }
%Besides the channels between two bodies, there is a single channel goes
%across the domain horizontally and the flow has been concentrated in
%this channel.  In the second frame of Fig.~\ref{fig:Eroding20vort}, this
%channel is between bodies 1 \& 20, 1 \& 9, 4 \& 3, 8 \& 11, 8 \& 2, 6 \&
%12, and 6 \& 19.  We can get a better understanding of this channel from
%the trajectories of tracers in Fig.~\ref{fig:Eroding20tracer}.  In Fig.~\ref{fig:Eroding20tracer}, 
%the flow (or tracer) first comes from three small channels between outer wall \& body
%10, bodies 7 \& 14, and bodies 14 \& 18, then the three branches become
%one and go into the main channel between bodies 1 \& 9.  From the first
%frame to the forth frame of Fig.~\ref{fig:Eroding20vort}, this channel
%is expanding faster than other channel and bodies 4 \& 12 has vanished
%and body 19 is almost gone in the forth frame.  The reason of fast
%expansion of this channel is the high vorticity.  In the forth frame of
%Fig.~\ref{fig:Eroding20vort}, we could see the vorticity is relatively
%high on the side of bodies which form this channel so the bodies are
%eroded fast on the side of this channel or have vanished completely.
%{\color{red} 
%Furthermore, how long the body lasts in the fluid depends on not only
%the size of the body but also the location of the body.  For example,
%the first three vanishing bodies are body 12 (as t=0.0256), 4 (as
%t=0.0278), then 17 (as t=0.0287).  Since body 4 is located in the main
%flow channel so the body 4 is vanished earlier than body 17 although the
%initial size of body 4 is larger than the size of body 17.  This channel
%will keep expanding and the bodies around this channel will vanish
%earlier than others.  After body 17, the next 8 vanished bodies with
%respect to time are body 19 (as t=0.0312),  body 7 (as t=0.0325), body
%20 (as t=0.0329), body 8 (as t=0.0349), body 10 (as t=0.0353), body 3
%(as t=0.0356), body 14 (as t=0.0363), then body 9 (as t=0.0376).  All of
%these eight bodies were around this channel before they are gone.
%
%For the flat faces at the sites of near contact, since we put the body
%be close to one another so we are able to see all the channels between
%two close bodies have flat faces.  Besides the channels between bodies,
%we also observe the flat faces on the bodies which are close to the
%outer wall. For example, we see the flat faces on the top of bodies 1 \&
%6 and the bottom of bodies 13 \& 15.  
%
%Since we use a Barycentric quadrature rule to approximate the velocity
%of tracers and apply the fourth order Runge Kutta method to find the new
%locations of tracers in each step, the tracers can be very close to but
%not penetrate the bodies.  Those body-closed tracers are moving
%perfectly so we are able to observe the details of the interaction of
%bodies and fluid.  Although the direction of flow is from left to right,
%the tracers will not only travel horizontally but also vertically.  In
%the bottom middle frame of Fig.~\ref{fig:Eroding20tracer}, some tracers
%climb up from right to left around the body.  There is a zoom windows of
%the climbing tracers trajectories in the bottom right frame of Fig.~\ref{fig:Eroding20tracer} and
%they are taken from the bottom middle frame of
%Fig.~\ref{fig:Eroding20tracer} with more tracers.  In the bottom right frame of 
%Fig.~\ref{fig:Eroding20tracer} the tracers around the center body first go
%from the channel below the body, climb up to the left, then go to the
%right and join another tracers in the channel above the body. 
%}
%\begin{figure}[H]
%\begin{center}
%\includegraphics[width = 0.32\textwidth]{./figs/tracer_20b210_zoom}
%\includegraphics[width = 0.32\textwidth]{./figs/tracer_20b240_zoom}
%\includegraphics[width = 0.32\textwidth]{./figs/tracer_20b270_zoom}
%\caption{\label{fig:Eroding20zoom}The zoom windows in the last three
%  frames of Figure~\ref{fig:Eroding20tracer}. The trajectories of
%  tracers around the bodies. Since we use a Barycentric quadrature rule
 % and the fourth order Runge Kutta method, the tracers can be very close
 % to but not penetrate the bodies.}
%\end{center}
%\end{figure}
%{\color{red}
%In Fig.~\ref{fig:Eroding20tort} we release 1000 tracers 
%from the starting cross section x = -1 to 
%the end cross section x = 1 and 
%use the local tortuosity to analysis the effect
%of the geometrical structure on fluid velocity field when the porosity is 62.9 \%.
%In Fig.~\ref{fig:Eroding20tort}(a) the graph represents the ratio of the
%x-component velocity u(-1, y) to its maximum velocity $u_{max}=2.9828
%\times 10^{-3}$ at equal spreading tracers on the starting cross section.
%The velocity is continuous along the starting cross section and
%acting as several Poiseuille flows. This velocity graph is different
%from the one in~\cite{matyka2008tortuosity} since we choose our cross
%section to be the beginning of the flow before the fluid touches any
%body. In Fig.~\ref{fig:Eroding20tort}(b) we record the local tortuosity
%$\tau(y)=\lambda(y)/d$
%of the tracers at $y$ on the starting cross section x = -1 
%where $d = 2$.
%This graph is not continuous 
%and the jumps in $\tau$ mean that the initial neighboring tracers have been
%seperated by the bodies. 
%The maximum of $\tau$ is 1.2713 and the minimum is 1.0001.
%}
%{\color{red}
%In Fig.~\ref{fig:Eroding20tort}(c) we calculate the difference of 
%$\tau$ at two neighboring tracers 
%and draw the trajectories of neighboring tracers
%if the difference of $\tau$ at the neighboring tracers is large. 
%From Fig.~\ref{fig:Eroding20tort}(c), we collect the 10 largest difference of $\tau$ 
%at the neighboring tracers and 
%draw the pair of trajectories in the same color and 
%are able to tell the neighboring trajectories will be very 
%different if they are separated by bodies.
%For each pair of trajectories, 
%they are splitted by the body at the left side of the body, then
%one of them goes up and the other one goes down along the body 
%until they rejoin with each other at the right side of body.  
%Furthermore, the pair trajectories could be splitted by more than one body, 
%for example in Fig.~\ref{fig:Eroding20tort}(c) the toppest red pair trajectories 
%are splitted by two close bodies.
%}

%To measure the local behavior of fluid quantitatively, 
%we use the velocities and local tortuosities at the starting cross section 
%to analysis the interaction of fluid and bodies until the end cross section 
%in Fig.~\ref{fig:Eroding100tort}(b) and (c).
%In Fig.~\ref{fig:Eroding100tort}(b) the graph represents the ratio of the
%x-component velocity u(-1, y) to its maximum velocity $u_{max}=3.9021
%\times 10^{-4}$ at equal spreading tracers on the strating cross section.
%As in the case of 20 nearly touching bodies, the velocity is continuous along
%the starting cross section and acting as a Poiseuille flow in the channel between
%bodies. When the fluid is close to bodies
%the speed of fluid will be slow down because of the no-slip condition on the bodies
%and this causes the separated branches of the inlet flow.
%The maximum velocity of the fluid with 100 bodies on the starting cross 
%section is slower than the maximum velocity of the fluid with 20 bodies
%but the velocity in the case of 100 bodies 
%changes more rapidly than the velocity of the fluid with 20 bodies.
%The reason of the rapid change on the starting cross section of 
%100-body case is because there are 
%more bodies near the starting cross section in
%100-body case than in the 20-body case. 
%so the inlet flow in 100-body case
%will be separated into more branches than the flow in 20-body case. 
%
%In Fig.~\ref{fig:Eroding100tort}(c) we record the 
%local tortuosity $\tau(y)$ for $y$ on the starting cross section x = -1.
%It is a discontinous function and the maximum of $\tau$ in the case of 100 bodies
%is 1.2547 and the minimum is 1.0003. This function $\tau$
%has more jumps than the one in the case of 20 bodies but its maximum 
%is smaller than the maximum in the case of 20 bodies. 
%The jump in this graph means the neighboring trajectories are splitted by bodies
%so it is reasonable to have more jumps in the case of 100-body than the case of 20-body
%since there are more bodies in the former case.

%To study the behavior of fluid in the porous media with multiple bodies, 
%we release 1000 tracers equispaced in $y$ from the 
%cross section x = -1 and record their trajectories and velocities
%until they arrive the end cross section x = 1 
%and the porosity of this geometric structure is 62.98 \%. 
%To understand the global action of fluid qualitatively, 
%we only draw 200 historical trajectories from x = -1 to x = 1 
%in Fig.~\ref{fig:Eroding100tort}(a). 
%When equispaced tracers are released in a 2D tube without any grains, 
%the streamlines are straight 
%and the intersection points of streamlines and any cross section are equispaced. 
%But it is not the same when there are bodies in the fluid. 
%In Fig.~\ref{fig:Eroding100tort}(a)
%even the tracers are equispaced in $y$ on the cross section $S$, 
%the streamlines have been push up and down by the bodies and
%some of them are close together. 
%
%
%%To understand the meaning of jump 
%%of $\tau$, we use the data we got from local tortuosity $\tau(y)$
%%to find the difference of $\tau$ at the neighboring tracers in Fig.~\ref{fig:Eroding100tort}(c) 
%%and draw 50 pairs of neighboring trajectories which have the largest difference of $\tau$
%%in Fig.~\ref{fig:Eroding100tort_traj}. These pairs of two neighboring trajectories are splitted by bodies
%%then rejoin with each other and the splitting-rejoining behavior causes the difference of $\tau$ with respect to the neighboring trajectories.   
%%We also can get the idea why the maximum value of $\tau$ is smaller 
%%than the one in the case of 20 bodies . 
%%The most of the bodies in this case are smaller than those in the case of 20 bodies.
%%So when two neighboring trajectories are splitted on the left side by a smaller body they do not
%%travel as far as the two splitted by a bigger body do after they rejoin on the right side.

%{\color{red}
%This case is 100 bodies in a Hagen-Poiseuille flow. The background
%flow is 
%\begin{align}
%  \UU(\xx)=U \left[
%  \begin{array}{c}
%    1-y^2 \\ 0
%  \end{array}
%  \right]
%\end{align}
%with floating flow rate $U$ such that pressure drop is fixed from x = -2 to x = 2.
%We record the vorticity of 
%fluid in colormap during the bodies erosion in Fig.~\ref{fig:Eroding100vort} and 
%measure the gap sizes between bodies in Fig.~\ref{fig:Eroding100gap} at four equispaced times.
%Since the vorticity is equivalent to the shear stress and the erosion is faster in the regions where
%the magnitude of the vorticity is larger 
%so the erosion is getting faster and faster 
%from the first snapshot to the forth one in Fig.~\ref{fig:Eroding100vort}.
%At t = 0.005 the porosity is 55.54\% and the most gap sizes between bodies are less than 0.015 in 
%Fig.~\ref{fig:Eroding100gap}(a). Since some of the bodies are vanished in the fluid, the numner of gaps 
%decreases and the gap sizes get larger and larger between the remaining bodies. 
%In Fig.~\ref{fig:Eroding100gap}(b) and (c) we still can see small gap sizes are in the majority but 
%it is not as concentrated as the one we see in Fig.~\ref{fig:Eroding100gap}(a). 
%In Fig.~\ref{fig:Eroding100gap}(d) there are only 254 gaps between 82 bodies and 
%the distribution of gap sizes spreads over without any major pack.  
%}
%
%\begin{figure}[H]
%\center
%\includegraphics*[width = 0.55\linewidth]{./figs/tort_eulerian}
%\caption{\label{fig:Eroding20tort_all} The tortuosity inside of an
%eroding geometry initialized with 20 bodies.  The tortuosity is
%calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue dots)
%and Lagrangian method~\eqref{eqn:tortuosity1} (red stars).  The red
%square correspond to the geometry in Figure~\ref{fig:Eroding20tort}(c).
%The dash line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$ as
%$p=0.2064$ and the root-mean-square error is $5.9 \times 10^{-3}$.}
%\end{figure}
%
%
%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/20b_second_moment_long_ref}
%\caption{\label{fig:Eroding20anomalousMedium} The analysis of anomalous
%dispersion in 20 nearly touching bodies with respect to initial porosity
%(as p=37.68\%) and another six porosities during the eroding. The dashed
%line correspondes to a porosity of 100\% and the dashed-dotted lines are
%lines of best fit with slope 1.0644 (as p=95.1\%), 1.0673 (as
%p=85.09\%), 1.0594 (as p=75.15\%), 0.9664 (as p=65.09\%), 0.7766 (as
%p=55.10\%), 0.7475 (as p=45.08\%), and 0.5607 (as p=37.68\%.}
%\end{figure}
%

%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/tort_eulerian20b}
%\caption{\label{fig:ErodingLow20tort} The tortuosity inside of an
%eroding geometry initialized with 20 bodies.  The tortuosity is
%calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue
%dots).  The dash line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$
%as $p=0.1669$ and the root mean square deviation is $1.13 \times
%10^{-2}$.}
%\end{figure}
%
%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/20b_dense_second_moment_ref}
%\caption{\label{fig:Eroding20anomalousLow} The analysis of anomalous
%dispersion in 20 nearly touching bodies with respect to initial porosity
%(as p=30.67\%) and another seven porosities during the eroding. The
%dashed line correspondes to a porosity of 100\% and the dashed-dotted
%lines are lines of best fit with slope 0.9194 (as p=95\%), 0.8666 (as
%p=85.17\%), 0.8682 (as p=75.10\%), 0.9087 (as p=65.02\%), 0.9054 (as
%p=55.08\%), 0.7234 (as p=45.03\%), 0.6895 (as p=35.05\%), and 0.9225 (as
%p=30.67\%).}
%\end{figure}
%
%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/tort_eulerian100}
%\caption{\label{fig:Eroding100tort_all} The Eulerian and Lagrangian
%approach of tortuosity.  The blue dots are Eulerian approach of
%tortuosity and the red stars and square are Lagrangian approach of
%tortuosity.  The dash line is the fitting line
%$\widehat{T}(\phi)=\phi^{-p}$ as $p=0.2459$ and the root-mean-square
%error is $5.5 \times 10^{-3}$.  We note that the model
%$\widehat{T}(\phi) = 1-p\ln(\phi)$, with $p=0.2631$ has a comparable
%root-mean-square error of $5.2 \times 10^{-3}$.}
%\end{figure}
%
%\begin{figure}
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/100b_second_moment_long_ref}
%\caption{\label{fig:Eroding100anomalous} The analysis of anomalous
%dispersion in 100 bodies with respect to initial porosity
%(as p=50.09\%) and another five porosities during the eroding. The dashed line
%correspondes to a porosity of 100\% and the dashed-dotted lines are lines
%of best fit with slope 1.1135 (as p=95.12\%), 0.6825 (as p=85.07\%), 
%0.7347 (as p=75.14\%), 0.5915 (as p=65.03\%), 
%0.6107 (as p=55.02\%), and 0.7 (as p=50.09\%).}
%\end{figure}
%
%\begin{figure}[H]
%\begin{center}
%\includegraphics[width = 0.8\textwidth]{./figs/100b_t100tracer}
%\caption{\label{fig:Eroding100tracer} The historical trajectories of 
%200 tracers released at x = -1 in the fluid with multiple bodies.}
%\end{center}
%\end{figure}

%\begin{figure}[H]
%\center
%\includegraphics*[width =0.5\linewidth]{./figs/tort100b_diff_top50}
%\caption{\label{fig:Eroding100tort_traj} The largest 50 difference of 
%local tortuosity $\tau$ at two neighboring tracers in
%100 bodies eroding when the porosity is 62.98 \%
%and the tortuosity is 1.1182. The same color of a pair of trajectories 
%means they are from two neighboring tracers.}
%\end{figure}

\end{document}