### 2019-10-23 docs/proposals/DOE/ErosionWhitePaper.tex
added new directory for potential NSF proposal

\documentclass[11pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage[sort,nocompress]{cite}
\usepackage{color}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{lastpage}
\usepackage{pgfplots}
\usepackage{rotating}
\usepackage{siunitx}
\usepackage{subfigure}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usepackage{titlesec}
\usepackage{todonotes}
\usepackage{wrapfig}
\usepackage{verbatim}
%\usepackage{showkeys}

\usetikzlibrary{shapes,arrows}

\addtolength{\oddsidemargin}{-0.75in}
\addtolength{\evensidemargin}{-0.75in}
\addtolength{\textwidth}{1.5in}
\addtolength{\topmargin}{-0.75in}
\addtolength{\textheight}{1.5in}
% For 11pt size

\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}

\newcommand{\bd}{\partial}
\newcommand{\BL}{\mathrm{BL}}
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\cc}{\mathbf{c}}
\newcommand{\dlp}{{\mathrm{dlp}}}
\newcommand{\ee}{\mathbf{e}}
\newcommand{\ff}{\mathbf{f}}
\newcommand{\FF}{\mathbf{F}}
\renewcommand{\gg}{\mathbf{g}}
\newcommand{\II}{\mathcal{I}}
\newcommand{\mcaption}[2]{\caption{\small \em #1}\label{#2}}
\newcommand{\nn}{\mathbf{n}}
\newcommand{\NN}{{\mathbf{N}}}
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\pderivtwo}[2]{\frac{\partial^{2} #1}{\partial #2^{2}}}
\newcommand{\pphi}{{\boldsymbol{\phi}}}
\renewcommand{\Re}{{\textrm{Re}}}
\newcommand{\rr}{{\mathbf{r}}}
\newcommand{\RR}{{\mathbb{R}}}
\newcommand{\slp}{{\mathrm{slp}}}
\newcommand{\ssigma}{{\boldsymbol{\sigma}}}
\newcommand{\SLP}{{\mathtt{SLP}}}
\newcommand{\DD}{{\mathcal{D}}}
\renewcommand{\SS}{{\mathcal{S}}}
\newcommand{\ttau}{{\boldsymbol{\tau}}}
\newcommand{\uu}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\VV}{{\mathcal{V}}}
\newcommand{\xx}{\mathbf{x}}
\newcommand{\yy}{\mathbf{y}}
\newcommand{\thL}{$\theta$--$L$}

\pagestyle{fancy}
\lhead{\footnotesize Bryan Quaife}
\chead{\footnotesize Viscous Erosion of a Porous Medium}
\rhead{\footnotesize \thepage}
\cfoot{}

%\usepgfplotslibrary{external}
%\tikzexternalize

\begin{document}
\begin{center}
Viscous Erosion of a Porous Medium \\
Bryan Quaife, Assistant Professor \\
Florida State University \\
512-436-1148, bquaife@fsu.edu \\
DOE/Office of Science Program Office: ASCR-Applied Mathematics \\
Funding Opportunity Announcement Number: DE-FOA-0001968
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{wrapfigure}[22]{r}{0.6\textwidth}
\centering
\includegraphics[width=0.58\textwidth]{figs/50bod.pdf}
\caption{\label{fig:50bod} \em Simulation of 50 bodies eroding in Stokes
flow under the action of fluid-mechanical erosion. The flow is
horizontal (left to right) and the 6 snapshots are evenly spaced in
time.  The local erosion rate on the bodies is proportional to the
magnitude of the vorticity which we compute in the fluid bulk (color
scale).  Erosion both diminishes the size of the bodies and considerably
alters their shapes.}
\end{wrapfigure}

Flow-induced erosion deteriorates and reshapes solid material over a
range of scales found in nature, including massive land formations
(mountains undergoing wind erosion) and watercourses (rivers and streams
undergoing hydraulic erosion).  Though less visible, erosion is working
at the very smallest scales, slowly deteriorating the individual
constituents of porous media (e.g.~soil, sand, or clay) or biological
structures like plaque and biofilms. Motivated by such examples, we are
interested in studying fluid-mechanical erosion in Stokes flow---the
most relevant regime for groundwater and bio-fluid applications.

In our recent publication~\cite{qua-moo2018}, we describe a suite of
numerical methods to simulate erosion in a viscous fluid
(Figure~\ref{fig:50bod}).  The highlights of our numerical methods are:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item {\bf Boundary Integral Equation Fluid Solver}: The
  fluid equations are solved with a boundary integral equation (BIE).
  The BIE formulation achieves high-order accuracy in complex geometries
  with optimal computational complexity.

  \item {\bf Regularizations of the Interface}: Erosion results in
  numerically challenging corners of the bodies.  We use a
  regularization term and a smoothing operation to control the
  smoothness of the moving interface.

  \item {\bf Interface Tracking}: We describe the interface using
  {\thL} coordinates, where $\theta$ is a local tangent angle, and $L$
  is the total interface length.  In these coordinates, the
  stiff terms are linear which allows implicit methods to be used for time discretization.

  \item {\bf Validation}: We compare the numerical simulations against analytical predictions for the limiting shape of a single body and scaling laws for its vanishing rate.
\end{enumerate}

\section{Proposed Work}
Having developed and validated numerical methods for simulating viscous
erosion, there are two main directions that our research will follow.
First, {\bf additional numerical tools are required to simulate more
realistic scenarios such as denser packings, transport, and
three-dimensional erosion}.  Second, {\bf we will investigate physically
motivated questions regarding the reconfiguration of the geometry and
transport inside an eroding porous medium}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved Numerical Methods}
\begin{wrapfigure}[20]{r}{0.35\textwidth}
\centering
\input{figs/vort1.tikz}
\caption{\label{fig:vort1} \em The vorticity along a vertical slice
cutting through a single body.  There are solid walls at $y = \pm 1$,
and an eroding body at $y = [-0.3,0.5]$.  The trapezoid rule gives large
errors near the solid walls, while the Barycentric quadrature formula
results in a smooth vorticity field.}
\end{wrapfigure}

Our current formulation fixes all bodies space.  However, in reality,
the bodies are also transported by the flow.  The PI is experienced with
simulating moving interfaces in Stokes flow~\cite{qua-bir2014,
qua-bir2016}. Recently, the PI and his former Ph.D.~student developed
time stepping methods for rigid body dynamics in Stokes
flow~\cite{bys-sha-qua2018}.  This work extends a contact
algorithm~\cite{lu-rah-zor2017} to simulate the dynamics of
nearly-touching rigid bodies.  The contact algorithm addresses time
stepping errors, but quadrature error introduced by the BIE formulation
must also be addressed.  A postdoc in the PI's group is extending a
Barycentric near-singular integration scheme~\cite{bar-wu-vee2015} to
develop quadrature formula to accurately compute the shear stress,
vorticity, and other physical quantities.  In Figure~\ref{fig:vort1},
the vorticity is computed along a vertical slice that cuts through a
single eroding body.  The trapezoid rule is accurate far from the solid
walls ($y = \pm 1$) and the boundary of the eroding body ($y \in
[-0.3,0.5]$), but introduces large errors near the walls and body.
However, our new Barycentric quadrature rule is uniformly accurate
throughout the entire fluid domain.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Physics-Based Understanding of Viscous Erosion}

Armed with these new tools, we plan to investigate how erosion alters a
porous medium through a reshaping process that occurs on the grain
scale. Initial investigations show that, as bodies erode and reduce in
size, the permeability of the medium increases. The rate of increase,
however, cannot be explained by the reduction in grain size alone, as
the reshaping and associated formation of channels are vital elements.
The new methods will allow us to examine these features in more detail.
In particular, we anticipate that erosion leads to the development of
{\bf anisotropic} permeability, as the medium gradually reconfigures to
transmit flow more easily in the direction of the initial flow. We will
characterize the changes in permeability and anisotropy especially in
nearly-touching body configurations, in more dense packings, and
eventually in configurations of mobile and sedimenting grains.

We will also investigate transport in eroding geometries. Two ways to
describe transport is with {\bf anomalous
diffusion}~\cite{dea-qua-bir-jua2018} and {\bf tortuosity}.  Anomalous
diffusion characterizes the spread of particles submerged in a geometry,
and the tortuosity quantifies the length of trajectories relative to
their size in a channel flow with no grains.  Both of these quantities
require a statistical analysis of streamlines, and we have already begun
these calculations (Figure~\ref{fig:streams}).  Once these quantities
are computed, we will investigate the relationship between erosion and
transport in eroding geometries.  The result will be new coarse-grained
descriptions for transport in complex eroding geometries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Biographical Information}

\begin{wrapfigure}[19]{l}{0.3\textwidth}
\centering
\includegraphics[width=0.28\textwidth]{figs/100b_t100tracer.pdf}
\caption{\label{fig:streams} \em Trajectories of passive particles in an
eroded geometry with 100 bodies.  The transport of flow in such a
geometry will be characterized by a statistical analysis of the
trajectories.}
\end{wrapfigure}
Since 2015, {\bf Bryan Quaife} has been an Assistant Professor in the
Department of Scientific Computing at Florida State University (FSU).
He is also a Research Associate in FSU's Geophysical Fluid Dynamics
Institute.  Dr.~Quaife received his Ph.D.~in 2011 in Applied and
Computational Mathematics at Simon Fraser University.  He was a
postdoctoral fellow from 2011--2015 in Dr.~George Biros' group at the
Institute for Computational Science and Engineering at the University of
Texas.  His research interests include the development of  high-fidelity
and efficient numerical methods for simulating fluid dynamics in complex
geometries.

The co-PI of this project is {\bf Nick Moore}.  Since 2014, Dr.~Moore
has been an Assistant Professor in the Department of Mathematics at
Florida State University.  He is also a Research Associate in FSU's
Geophysical Fluid Dynamics Institute.  Dr.~Moore received his Ph.D.~in
2010 in Mathematics at the University of North Carolina, Chapel Hill. He
was a postdoctoral researcher at the Courant Institute of Mathematical
Sciences from 2010--2014 working with Dr.~Michael Shelley and others.
His expertise includes modeling dynamic interactions between fluids and
structures, especially in geophysically and biologically motivated
settings.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Expected Budget}
We request funding for three years.  For the project to be successful,
we propose one month of summer salary each year for both the PI and
co-PI, one joint two-year postdoctoral fellow, two full-time RAs (one
from each of our home departments), laptops, and travel money.
Including FSU's overhead, we expect that {\bf the total cost of the
project would be around \$750,000.}
%\begin{itemize}
%  \item Summer salary (1 month a year for 3 years)
%  \item 2 RA (1 in Math, 1 on DSC)
%  \item 1 joint postdoc
%  \item Travel
%  \item Computers
%  \item FSU overhead
%\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\footnotesize{\bibliography{refs}}

\end{document}