### 2019-11-14 docs/proposals/NSF/ProjDocs/facilities.tex

%\documentclass[12pt]{amsart}
\documentclass[11pt]{article}

\usepackage{color}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{mathpazo}

%\renewcommand\thesubsection{(\alph{subsection})}
%\renewcommand\thesubsubsection{(\roman{subsubsection})}

\addtolength{\oddsidemargin}{-0.75in}
\addtolength{\evensidemargin}{-0.75in}
\addtolength{\textwidth}{1.5in}
\addtolength{\topmargin}{-0.75in}
\addtolength{\textheight}{1.5in}
% For 11pt size

\newcommand{\todo}[1]{ \fbox{{\bf TODO:} \color{red} #1}}

\pagestyle{fancy}
\lhead{\footnotesize Quaife and Moore}
\chead{\footnotesize Facilities, Equipment, and Other Resources}
\rhead{\footnotesize \thepage{ of }\pageref{LastPage}}
\cfoot{}

\begin{document}
\thispagestyle{fancy}

\subsection*{Computing Resources}
Florida State University's (FSU) high performance computing (HPC) and
high throughput computing (HTC) clusters are managed by the Research
Computing Center (RCC).

The FSU HPC system is comprised of 12,492 x86 64-bit compute cores
linked together by low-latency infiniband networks for MPI
communication. The aggregate peak performance of the system is 264.6
TFLOPS. Compute nodes support between 2 and 16GB of memory per core. A
redundant cluster of specialized nodes serve as the user entry points
for the system. The Slurm job scheduler handles job submission and
scheduling on the HPC system. A broad set of compilers, math and
communication libraries, and software applications are made available on
the HPC to maximize the utility of the system by users across diverse
academic disciplines.  The cluster also includes nodes with 16 NVIDIA
Tesla gpGPU cards (m2050) for a total 7,168 GPU cores for a peak system
performance of 19.8 TFLOPS.

The FSU HTC system, Condor, is comprised of 1,200 x86 64-bit compute
cores.  The system is specifically tailored to support large batches of
serial jobs where overall throughput is preferred over performance.
The aggregate peak performance of the system is 16.5 TFLOPS.  Compute
nodes support between 2 and 8 GBs of memory per core.

In addition to the HPC and HTC systems, the PI's group has access to
RCC's Spear cluster.  The Spear Cluster is comprised of 288 x86 64-bit
process cores linked together by a QDR Infiniband network.  Eight of the
Spear nodes also connect to a PCI chassis equipped with 16 Nvidia Tesla
gpGPU cards (M2050) for a total of 7,168 GPU cores for a peak system
performance of 19.8 TFLOPS.

\subsection*{Personnel}
The RCC has dedicated personnel who offer help with hardware
specification and acquisition, software acquisition and licensing,
compilation, installation, integration with the queuing system, running
benchmarks, developing computational workflows, and necessary end-user
training.  In addition, the Department of Mathematics and the Department
of Scientific Computing have systems administrators that assist with the
PIs's group's computing needs.

Both the PIs are associates in the Geophysical Fluid Dynamics Institute
(GFDI). GFDI is a fluids lab where experimental work will be performed.
The lab includes the necessary support staff, including a machinist and
system administrator, to assist with experimental setups and
computational needs.

\subsection*{Office Space}
The project will utilize office and workspace in the Department of
Mathematics, Department of Scientific Computing, and Geophysical Fluid
Dynamics Institute at Florida State University. Offices are equipped
with a standard desktop computer and internet access.

\end{document}