### 2019-11-19 docs/proposals/NSF/ProjDocs/funding_BQ.tex

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{fancyhdr}
\usepackage{lastpage}
%\usepackage{amsmath,amsfonts,amssymb}
\usepackage{tabularx}
\usepackage{todonotes}
\usepackage{mathpazo}

%\renewcommand\thesubsection{(\alph{subsection})}
%\renewcommand\thesubsubsection{(\roman{subsubsection})}

\addtolength{\oddsidemargin}{-0.75in}
\addtolength{\evensidemargin}{-0.75in}
\addtolength{\textwidth}{1.5in}
\addtolength{\topmargin}{-0.75in}
\addtolength{\textheight}{1.5in}
% For 11pt size

\pagestyle{fancy}
\lhead{\footnotesize Bryan Quaife}
\chead{\footnotesize Current and Pending Support}
\rhead{\footnotesize \thepage{ of }\pageref{LastPage}}
\cfoot{}

\begin{document}

\subsection*{Current Support:}
{\bf Title}: Computational Methods for Complex Stokesian Fluids \\
{\bf Source}: Simons Foundation, Mathematics and Physical
Sciences-Collaboration Grants for Mathematicians \\
{\bf Total Award Amount:} \$25,200 (Travel funds only) \\
{\bf Duration:} September 1, 2017--August 31, 2020 \\
{\bf Person-Months Per Year Committed to Project:} Calendar: 0,
Academic: 0, Summer: 0 \\

\noindent
{\bf Title}: The Role of Vorticity and Fuel Moisture on the Near-Field
Plume Structure and Ember Dynamics \\
{\bf Source}: Department of Defense, Strategic Environmental Research
and Development Program \\
{\bf Total Award Amount:} \$2,235,917 \\
{\bf Duration:} May 2020--January 2024 (This is a collaborative grant
that supports 7 PIs from 4 institutions) \\
{\bf Person-Months Per Year Committed to Project:} Calendar: 0,
Academic: 0, Summer: 1 \\

\vspace{-17pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Pending Support:}
{\bf Title}: Erosion, Transport, and Dispersion in Granular and Porous
Media (this proposal) \\
{\bf Source}: National Science Foundation, Division of Mathematical
Sciences, Computational Mathematics \\
{\bf Total Award Amount:} \$383,086 \\
{\bf Duration:} July 1, 2020--June 30, 2023 \\
{\bf Person-Months Per Year Committed to Project:} Calendar: 0,
Academic: 0, Summer: 1 \\

\end{document}