### 2020-07-29 docs/proposals/NSF/PostDocs/PI_change_request.tex
added letter requesting change of PI

\documentclass{article}

\usepackage[dvinames]{xcolor}
% Measurements are taken directly from the guide
\usepackage[top=2in,left=1.5in,bottom=0.5in,right=0.625in]{geometry}
\usepackage{graphicx}
\usepackage[colorlinks=false,
pdfborder={0 0 0},
]{hyperref}
\usepackage[absolute]{textpos}
\usepackage{tikz}
\usetikzlibrary{calc}
% A nice serif font, but no the prescribed
% nonfree ITC stone
\usepackage[oldstylenums]{kpfonts}
\usepackage[T1]{fontenc}

\definecolor{fsugarnet}{RGB}{120,47,64}
\definecolor{fsugold}{RGB}{206,184,136}
\definecolor{fsublack}{RGB}{44,42,41}
\definecolor{fsuwhite}{RGB}{255,255,255}

% No paragraph indentation
\parindent0pt
\setlength{\parskip}{0.8\baselineskip}
\raggedright
\pagestyle{empty}
% Ensure consistency in the footer
\urlstyle{sf}

\begin{document}

\today
\bigskip

Dear Dr.~Peszynska:

I would like to request a restructuring of the NSF grant with Award
Number 2012560. PI Moore has left Florida State University (FSU) and
taken a faculty position at the US Naval Academy. Dr.~Moore's new
department does not have a graduate program, so the funded graduate
student's senior advisor will be co-PI Quaife at FSU. So that the grant
can be managed by FSU, I request that Dr.~Quaife becomes the lead PI and
Dr.~Moore becomes the co-PI. I also request permission for FSU to create
a sub-contract for the US Naval Academy that will manage Dr.~Moore's
funding.

This request is written in consultation with Dr.~Moore. Please contact
me if further clarification is required.

Sincerely,
\vspace{2\baselineskip}

Bryan Quaife \\
Assistant Professor

% -------------------------------------------------------
% Add logo, the text under the crimson line,
% and the line itself
\begin{textblock*}{2in}[0.3066,0.39](1.5in,1.33in)
\includegraphics[width=2in]{fsu_logo}
\end{textblock*}
\begin{textblock*}{6.375in}(1.5in,1.55in)
% 6.375=8.5 - 1.5 - 0.625
\sffamily
\hfill \color{fsugarnet} Department of Scientific Computing \\
\hfill Dr.~Bryan Quaife,
Assistant Professor
\end{textblock*}
\begin{tikzpicture}[remember picture,overlay]
\draw[color=fsugarnet,line width=1pt] (current page.north west)+(1.5in,-1.5in) -- ($(-0.625in,-1.5in)+(current page.north east)$);
\end{tikzpicture}

% Add the footnote. A
% better way would be to
% use fancyhdr
\vfill
{\footnotesize\color{fsugarnet}\sffamily
400 Dirac Science Library, Tallahassee, FL
32306-4120 \\
[-0.1\baselineskip]
(512) 436-1148 \textbullet\
\url{bquaife@fsu.edu}
\textbullet\
\url{http://www.sc.fsu.edu/~bquaife}
}
\end{document}