### 2020-07-29 docs/proposals/NSF/budgets/budgetImpact.tex

\documentclass[11pt]{article}
\usepackage{fullpage}
\topmargin0.25in

\begin{document}
\thispagestyle{empty}
\sloppy

\begin{center}
\Large \bf Budget Impact Statement
\end{center}
\vspace{0.2in}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The PI and co-PI have successful reduced the budget for this project
from the originally requested \$$439,192$ to under \$$250,000$. The
reduction will have some impact on the scope of the project, but we are
confident that we will still be able to meet the main goals of the
project. First, the largest change is that instead of (partially)
supporting two graduate students for the 3 years, only one graduate
student will be supported. This graduate student will spearhead the
implementation of novel numerical algorithms for simulating flow-induced
erosion with multi-physical effects to be used in the modeling of dense
suspensions, groundwater systems, and sinkhole collapse. The graduate
student will also explore machine learning techniques in order to inform
reduced models that learn from batches of high-fidelity simulations.
While we still expect this machine learning component to be carried out,
there may be some reduction in its extent due to the removal of the
second graduate student. Thus, the main computational thrust of the
project will remain largely intact, with possible modifications only in
the depth and extent of each component, not in the components
themselves.

Notably, we have removed all expenses pertaining to the laboratory
experiments that were originally intended to corroborate and interplay
with the numerical simulations. The cuts include the second graduate
student (who would have spent part of their time interacting with the
experiments), the funding for an undergraduate student, the funding for
the technical person, and a large portion of the materials and supplies.
While the laboratory component would have been a nice addition to the
project, its role is not as central to the project as the computational
component is. Further, this experimental component could possibly be
replaced by a collaboration with other, external experimentalists.

In addition, we have cut 1/2 month of summer salary support for co-PI
Quaife. We have retained the originally requested summer salary for PI
Moore since, with the cut of the second graduate student, the PI will
need to take on some additional burdens of the project. We have also
significantly scaled back the travel support as COVID-19 may likely
curtail the ability to travel anyway and many conferences will likely be
held remotely.

\end{document}