### 2020-08-07 docs/proposals/NSF/PostDocs/Scope_of_work_FSU.tex
starting scope of work

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{fancyhdr}
\usepackage{lastpage}
%\usepackage{amsmath,amsfonts,amssymb}
\usepackage{tabularx}
\usepackage{todonotes}
\usepackage{mathpazo}

%\renewcommand\thesubsection{(\alph{subsection})}
%\renewcommand\thesubsubsection{(\roman{subsubsection})}

\addtolength{\oddsidemargin}{-0.75in}
\addtolength{\evensidemargin}{-0.75in}
\addtolength{\textwidth}{1.5in}
\addtolength{\topmargin}{-0.75in}
\addtolength{\textheight}{1.5in}
% For 11pt size

\pagestyle{fancy}
\lhead{\footnotesize Bryan D.~Quaife}
\chead{\footnotesize FSU Scope of Work}
\rhead{\footnotesize \thepage{ of }\pageref{LastPage}}
\cfoot{}

\begin{document}
\section*{Personnel}
The research at FSU will be performed by PI Quaife and Ph.D.~student
Jake Cherry for the duration of the award.

\section*{Deliverables}
\paragraph{Simulating mechanical and chemical erosion, as well as
transport, including sedimentation and compaction, of eroding bodies:}

\paragraph{Simulating the transport of finite Peclet number
concentrations in eroded porous medium.}

\paragraph{Improving numerical methods including fast summation methods,
preconditioners, and quadrature.}

\section*{Timeline}

\end{document}