### 2022-10-13 docs/proposals/NSF/budgets/budgetJust_Colgate.tex
NM

\documentclass[11pt]{article}
\usepackage{fullpage}
\topmargin0.25in

\begin{document}
\thispagestyle{empty}
\sloppy

\begin{center}
\Large \bf Budget Justification
\end{center}
\vspace{0.2in}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{A: Senior Personnel (co-PI: Matthew Moore)}
%\hfill \\
%\noindent
The requested senior personnel funding is to support one month of 
co-PI Moore's summer salary in year 3. During this period, the PIs
will work together to develop and implement algorithms, write papers,
and work closely with the graduate student. 
Co-PI Moore's summer salary will in year 3 will be \$93,788.
with the monthly salary listed in the table:
%\hfill \\
\begin{center}
  \begin{tabular}{|c|r|r|}
    \hline
    & Year 3  & Total \\
    \hline
    co-PI Moore & $\$10,421$  & $\$10,421$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{C: Fringe Benefits}
The fringe benefits of co-PI Moore at Colgate University are 32\%.  \\
%In addition a COLA of 3\% is added to estimate the cost of living increase government works may receive in January. \\
%\hfill \\
\begin{center}
  \begin{tabular}{|c|r|r|}
    \hline
    & Year 3 & Total \\
    \hline
    co-PI Moore Fringe & $\$3,335$ & $\$3,335$ \\
    \hline
    %co-PI Moore COLA & $\$288$ & $\$293$ & $\$299$ & $\$880$ \\
    \hline
    Total & $\$3,335$ & $\$3,335$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{E: Domestic Travel}
%\hfill \\
%\noindent
The requested travel funds is to partially fund a domestic trip
for co-PI Moore in year 3.  The expected costs per
conference participant is $\$1,000$ with a breakdown of $\$300$ for
airfare, $\$300$ for lodging, $\$56$ for local transportation, $\$144$
for meals (four days at $\$36$ per day), and $\$200$ for conference
registration.  Potential meeting venues include several SIAM Meetings
(Annual Meetings, CSE, Dynamical Systems), APS Fluids Meetings, or
appropriately themed workshops.  Alternatively, co-PI Moore may use
these travel funds to visit PI Quaife and the funded graduate student at
FSU in order to collaborate on the project.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    & Year 3 & Total \\
    \hline
    Domestic Travel & $\$1,000$ & $\$1,000$ \\
    \hline
  \end{tabular}
\end{center}

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection*{G: Other Direct Costs}
%%\noindent
%\paragraph{\bf Materials and Supplies:} The requested materials and
%supplies of $\$1000$ in year 1 is to partially fund a laptop or desktop
%computer for co-PI Moore. The computer will be used for developing the
%proposed algorithms and presenting the results at conferences.
%\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{H: Total Direct Costs}
%\hfill \\
%\noindent
Total of Sections A--G.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|}
    \hline
    Year 3 & Total \\
    \hline
    $\$14,756$ & $\$14,756$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{I: Indirect Costs}
%\hfill \\
%\noindent
Colgate University's indirect cost rate is 35\% for research on
modified total direct costs (MTDC).  The base for overhead is $\$14,756$ in
year 3.  
%The US Naval Academy defines a fiscal year to be October 1-September 30.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|}
    \hline
    Year 3 & Total \\
    \hline
    $\$5,165$ & $\$5,165$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{J: Total Direct and Indirect Costs}
%\hfill \\
%\noindent
Total of Sections H and I.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|}
    \hline
    Year 3 & Total \\
    \hline
    $\$19,921$ & $\$19,921$ \\
    \hline
  \end{tabular}
\end{center}

\end{document}