### 2022-12-02 docs/proposals/NSF/PostDocs/Colgate_Transfer/budgetJustificationFSU2022.tex
working on revisions

\documentclass[11pt]{article}
\usepackage{fullpage}
\topmargin0.25in

\begin{document}
\thispagestyle{empty}
\sloppy

\begin{center}
\Large \bf Budget Justification
\end{center}
\vspace{0.2in}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{A: Senior Personnel (PI: Bryan Quaife)}
%\hfill \\
%\noindent
The requested senior personnel funding is to support one month of the
PI's summer salary in years 1 and 2, and 0.25 months of the PI's summer
salary in year 3. During this period, the PI will work together to
develop and implement algorithms, write papers, and work closely with
subcontractor Moore and the graduate student. The PI's monthly salary in
year one is $\$10,553$. For years 2 and 3, a 3\% increase is included.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|r|r|r|r|}
    \hline
    & Year 1 & Year 2 & Year 3  & Total \\
    \hline
    PI & $\$10,553$ & $\$10,869$ & $\$2,799$ & $\$24,221$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{B: Other Personnel}
%\hfill \\
%\noindent
The requested other personnel funding includes salary for one full-time
Ph.D.~student in two semesters of years one, two, and three, starting at
$\$21,000$. The Ph.D.~student will develop and perform numerical
simulations. An annual increase of 3\% is included in years 2 and 3.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    & Year 1 & Year 2 & Year 3 & Total \\
    \hline
    Ph.D. & $\$21,000$ & $\$21,630$ & $\$22,279$ & $\$64,909$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{C: Fringe Benefits}
{\bf PI:} The fringe benefits of the PI are 18.54\%. \\
{\bf Ph.D.:} The fringe benefits of the graduate student are
0.6\% plus an $\$1,965$ health insurance. \\
%\hfill \\
\begin{center}
  \begin{tabular}{|c|r|r|r|r|}
    \hline
    & Year 1 & Year 2 & Year 3 & Total \\
    \hline
    PI & $\$1,956$ & $\$2,015$ & $\$519$ & $\$4,490$ \\
    \hline
    Ph.D. & $\$2,091$ & $\$2,095$ & $\$2,099$ & $\$6,285$ \\
    \hline
    Total & $\$4,047$ & $\$4,110$ & $\$2,618$ & $\$10,775$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{E: Domestic Travel}
%\hfill \\
%\noindent
The requested travel funds is to fund an annual domestic trip for the PI
or the Ph.D.~student in years 2 and 3. The expected costs per conference
participant is $\$1,097$ with a breakdown of $\$397$ for airfare,
$\$300$ for lodging, $\$56$ for local transportation, $\$144$ for meals
(four days at $\$36$ per day), and $\$200$ for conference registration.
An additional $\$523$ is budgeted for year 3. Potential meeting venues
include several SIAM Meetings (Annual Meetings, CSE, Dynamical Systems),
APS Fluids Meetings, or appropriately themed workshops. The PI and
Ph.D.~student will give presentations and actively participate in the
meeting or workshop. Travel funds for the first year are removed due to
the COVID crisis.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    & Year 1 & Year 2 & Year 3 & Total \\
    \hline
    Domestic Conferences &  & $\$2,195$ & $\$2,718$ & $\$4,813$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{G: Other Direct Costs}
%\noindent
\paragraph{\bf Materials and Supplies:} The requested materials and
supplies is to purchase one $\$1,359$ laptop with accessories, which is
not otherwise available, for the PI and Ph.D.~student in year 1. The
laptops will be used for developing the proposed algorithms and
presenting the results at conferences. The laptop will give the group
members the flexibility to work on the proposed research remotely. The
laptop will be used exclusively for the proposed research. 
\\
\paragraph{\bf Subawards:} The requested subaward is for Dr.~Nick Moore,
previously employed at the US Naval Academy (USNA), and currently
employed at Colgate University, to provide two of the deliverables as
described in the scope of work.
\\
\paragraph{\bf Tuition:} The requested other direct costs are to cover the
Ph.D.~student's tuition for 18 credit hours, at a rate of $\$407.55$ per
credit hour in year 1. An annual increase of 1\% is included per
published university projections.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    & Year 1 & Year 2 & Year 3 & Total \\
    \hline
    Materials and Supplies & $\$1,500$ & & & $\$1,500$ \\
    \hline
    Subawards & $\$14,445$ & $\$15,811$ & $\$19,921$ & $\$50,177$ \\
    \hline
    Subaward Location 
    & USNA & USNA & Colgate & \\
    \hline
    Tuition & $\$7,336$ & $\$7,409$ & $\$7,484$ & $\$22,229$ \\
    \hline
    Total & $\$23,281$ & $\$23,220$ & $\$27,405$ & $\$73,906$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{H: Total Direct Costs}
%\hfill \\
%\noindent
Total of Sections A--G.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|}
    \hline
    Year 1 & Year 2 & Year 3 & Total \\
    \hline
    $\$58,881$ & $\$62,024$ & $\$57,819$ & $\$178,724$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{I: Indirect Costs}
%\hfill \\
%\noindent
Florida State University's federally negotiated indirect cost rate is
54\% for research on modified total direct costs (MTDC) which excludes
tuition subaward costs after the first \$25,000. The base for overhead
is $\$51,545$ in year 1, $\$49,359$ in year 2, and $\$30,414$ in year 3.
Florida State University defines a fiscal year to be July 1--June 30.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|}
    \hline
    Year 1 & Year 2 & Year 3 & Total \\
    \hline
    $\$27,834$ & $\$26,654$ & $\$16,424$ & $\$70,912$ \\
    \hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{J: Total Direct and Indirect Costs}
%\hfill \\
%\noindent
Total of Sections H and I.
%\hfill \\
\begin{center}
  \begin{tabular}{|c|c|c|c|}
    \hline
    Year 1 & Year 2 & Year 3 & Total \\
    \hline
    $\$86,715$ & $\$88,678$ & $\$74,243$ & $\$249,636$ \\
    \hline
  \end{tabular}
\end{center}

\end{document}